---
title: "Extended Kinome Network v2 (Conservative) — Build Pipeline"
subtitle: "OmniPath + HIPPIE expansion (HPA filter retained for Cantley)"
author: "Chinmaya Joisa"
date: today
toc: true
toc-depth: 3
toc-title: Table of Contents
format:
  html:
    page-layout: full
    embed-resources: true
    code-fold: true
    code-tools: true
execute:
  echo: true
  cache: false
  message: false
  warning: false
  fig-align: center
  fig-dpi: 300
---

```{r setup, include=FALSE}
require("knitr")
knitr::opts_knit$set(root.dir = here::here())
set.seed(123)
```

```{r libraries}
library(data.table)
library(tidyverse)
library(fst)
library(here)
library(scales)
library(ggpubr)
library(patchwork)
library(DT)
library(clusterProfiler)
library(org.Hs.eg.db)
library(OmnipathR)
```

# Overview

This is the **conservative** v2 build strategy:
- Keep Cantley with **HPA co-localization filter** (same as v1)
- Add **OmniPath** enzyme-substrate layer (high-quality curated)
- Add **HIPPIE** substrate-substrate PPIs (high-confidence interactions)
- Add **HIPPIE brute-force** for missing drug targets (connectivity-based recovery)

| Layer | Treatment |
|---|---|
| Curated KS pairs | Unchanged from v1 |
| Cantley motif + promiscuity | Same thresholds (95th percentile, 50th promiscuity) |
| Cantley HPA co-localization | **RETAINED** (same as v1) |
| Cantley Ochoa functional score | Same threshold (75th percentile) |
| OmniPath | NEW (PhosphoSite + SIGNOR curated) |
| HIPPIE substrate-substrate | NEW (high-confidence PPIs) |
| HIPPIE drug target expansion | NEW (brute-force recovery) |

# Step 1: Load v1 Combined Pairs (with HPA filter)

```{r load_v1_pairs}
combined_v1 <- read_csv(here("results/combined_kinase_substrate_pairs_2025.csv"),
                        show_col_types = FALSE) %>%
  mutate(Kinase_entrez = as.character(Kinase_entrez),
         Substrate_entrez = as.character(Substrate_entrez))

cat("v1 Combined KS pairs (HPA-filtered):\n")
cat("  Total pairs:          ", nrow(combined_v1), "\n")
cat("  From Cantley:         ", sum(combined_v1$Source == "Cantley"), "\n")
cat("  From curated sources: ", sum(combined_v1$Source != "Cantley"), "\n")
cat("  Unique substrates:    ", n_distinct(combined_v1$Substrate_entrez), "\n")
cat("  Unique kinases:       ", n_distinct(combined_v1$Kinase_entrez), "\n")
```

# Step 2: Add OmniPath Enzyme-Substrate Layer

```{r omnipath_layer}
cat("Querying OmniPath (PhosphoSite + SIGNOR)...\n")
omnipath_es <- enzyme_substrate(
  organism   = 9606,
  resources  = c("PhosphoSite", "SIGNOR"),
  genesymbols = TRUE,
  silent     = TRUE
)

curated_kinases <- unique(combined_v1$Kinase)

omnipath_sub_symbols <- omnipath_es %>%
  dplyr::filter(enzyme_genesymbol %in% curated_kinases) %>%
  dplyr::filter(residue_type %in% c("S", "T", "Y")) %>%
  dplyr::select(enzyme_genesymbol, substrate_genesymbol, residue_type, residue_offset) %>%
  distinct()

cat("OmniPath records (curated kinases + S/T/Y):", nrow(omnipath_sub_symbols), "\n")

# Map substrate gene symbols → Entrez IDs
omnipath_entrez_map <- bitr(
  unique(omnipath_sub_symbols$substrate_genesymbol),
  fromType = "SYMBOL", toType = "ENTREZID",
  OrgDb = org.Hs.eg.db, drop = TRUE
) %>%
  dplyr::rename(substrate_genesymbol = SYMBOL, Substrate_entrez = ENTREZID) %>%
  distinct()

# Map enzyme gene symbols → Entrez IDs
omnipath_kin_map <- bitr(
  unique(omnipath_sub_symbols$enzyme_genesymbol),
  fromType = "SYMBOL", toType = "ENTREZID",
  OrgDb = org.Hs.eg.db, drop = TRUE
) %>%
  dplyr::rename(enzyme_genesymbol = SYMBOL, Kinase_entrez = ENTREZID) %>%
  distinct()

omnipath_pairs <- omnipath_sub_symbols %>%
  left_join(omnipath_entrez_map, by = "substrate_genesymbol") %>%
  left_join(omnipath_kin_map,    by = "enzyme_genesymbol") %>%
  dplyr::filter(!is.na(Substrate_entrez), !is.na(Kinase_entrez)) %>%
  transmute(
    Kinase           = enzyme_genesymbol,
    Kinase_entrez    = as.character(Kinase_entrez),
    Substrate        = substrate_genesymbol,
    Substrate_entrez = as.character(Substrate_entrez),
    Site             = paste0(residue_type, residue_offset),
    Curated_site     = TRUE,
    Likely_functional = NA,
    Evidence         = "OmniPath enzyme-substrate (PhosphoSite/SIGNOR curated)",
    Reference_PMID   = NA_character_,
    Source           = "OmniPath"
  ) %>%
  distinct()

cat("OmniPath pairs after Entrez mapping:", nrow(omnipath_pairs), "\n")
cat("Unique OmniPath substrates:", n_distinct(omnipath_pairs$Substrate_entrez), "\n")

# How many are novel (not already in v1)?
novel_omnipath_subs <- setdiff(omnipath_pairs$Substrate_entrez, combined_v1$Substrate_entrez)
cat("Novel substrates added by OmniPath:", length(novel_omnipath_subs), "\n")

# Bind to v1 pairs
combined_with_omnipath <- bind_rows(combined_v1, omnipath_pairs) %>% distinct()
cat("Total combined pairs after OmniPath:", nrow(combined_with_omnipath), "\n")
```

# Step 3: Add HIPPIE Substrate-Substrate PPIs

```{r hippie_substrate_substrate}
cat("Loading HIPPIE network...\n")
hippie <- fread(
  here("data/hippie_current.txt"),
  col.names = c("Gene1", "Entrez1", "Gene2", "Entrez2", "Score", "Metadata"),
  colClasses = list(character = c("Entrez1", "Entrez2"))
)

cat("HIPPIE interactions:", nrow(hippie), "\n")
cat("Unique proteins:    ", uniqueN(c(hippie$Entrez1, hippie$Entrez2)), "\n")

# Define substrate and kinase sets
substrate_ids <- unique(combined_with_omnipath$Substrate_entrez)
kinase_ids     <- unique(combined_with_omnipath$Kinase_entrez)
substrate_only <- setdiff(substrate_ids, kinase_ids)

# Confidence cutoff (top 25%)
confidence_cutoff <- quantile(hippie$Score, 0.75, na.rm = TRUE)
cat("HIPPIE confidence cutoff (75th pctile):", round(confidence_cutoff, 3), "\n")

# Substrate-substrate PPIs (both proteins are substrates-only)
hippie_ss <- hippie[
  Score > confidence_cutoff &
    Entrez1 %in% substrate_only & Entrez2 %in% substrate_only &
    Entrez1 != Entrez2
]

# Deduplicate undirected edges
hippie_ss_dedup <- hippie_ss[
  , .(p1 = as.character(pmin(as.integer(Entrez1), as.integer(Entrez2))),
      p2 = as.character(pmax(as.integer(Entrez1), as.integer(Entrez2))),
      Gene1, Gene2, Score)
] %>% unique(by = c("p1", "p2"))

cat("HIPPIE substrate-substrate edges (high-confidence):", nrow(hippie_ss_dedup), "\n")
cat("Unique substrates touched:", uniqueN(c(hippie_ss_dedup$p1, hippie_ss_dedup$p2)), "\n")

# Build symbol maps
hippie_sym1 <- bitr(hippie_ss_dedup$p1, fromType = "ENTREZID", toType = "SYMBOL",
                    OrgDb = org.Hs.eg.db, drop = TRUE) %>%
  dplyr::rename(p1 = ENTREZID, Protein1 = SYMBOL) %>% distinct()
hippie_sym2 <- bitr(hippie_ss_dedup$p2, fromType = "ENTREZID", toType = "SYMBOL",
                    OrgDb = org.Hs.eg.db, drop = TRUE) %>%
  dplyr::rename(p2 = ENTREZID, Protein2 = SYMBOL) %>% distinct()

hippie_ss_final <- hippie_ss_dedup %>%
  left_join(hippie_sym1, by = "p1") %>%
  left_join(hippie_sym2, by = "p2") %>%
  dplyr::filter(!is.na(Protein1), !is.na(Protein2)) %>%
  transmute(
    Protein1, Protein2,
    Protein1_entrez = p1,
    Protein2_entrez = p2,
    Evidence = "protein-protein interaction",
    Source   = "HIPPIE"
  ) %>%
  distinct()

cat("HIPPIE substrate-substrate pairs (with symbols):", nrow(hippie_ss_final), "\n")
```

# Step 4: HIPPIE Brute-Force for Missing Drug Targets

```{r hippie_brute_force}
# Load drug target panel
drug_targets <- read.fst(here("data/kinome_profiling_novartis_evebio_combined.fst")) %>%
  as_tibble() %>%
  distinct(target, target_entrez_id) %>%
  mutate(target_entrez_id = as.character(target_entrez_id))

# Current network node set (after KS + OmniPath + HIPPIE SS)
ks_nodes <- union(
  as.character(combined_with_omnipath$Kinase_entrez),
  as.character(combined_with_omnipath$Substrate_entrez)
)
all_current_nodes <- union(ks_nodes, c(hippie_ss_final$Protein1_entrez, hippie_ss_final$Protein2_entrez))

missing_drug_targets <- drug_targets %>%
  dplyr::filter(!target_entrez_id %in% all_current_nodes)

cat("Drug targets not yet in network:", nrow(missing_drug_targets), "/", nrow(drug_targets), "\n")

# Find HIPPIE edges connecting missing targets to in-network proteins
hippie_bf <- hippie[Score > confidence_cutoff] %>%
  dplyr::filter(
    (Entrez1 %in% missing_drug_targets$target_entrez_id & Entrez2 %in% all_current_nodes) |
    (Entrez2 %in% missing_drug_targets$target_entrez_id & Entrez1 %in% all_current_nodes)
  ) %>%
  mutate(
    # Orient: in-network protein is Protein1, missing target is Protein2
    Protein1_entrez = case_when(
      Entrez1 %in% all_current_nodes ~ Entrez1,
      TRUE ~ Entrez2
    ),
    Protein2_entrez = case_when(
      Entrez1 %in% missing_drug_targets$target_entrez_id ~ Entrez1,
      TRUE ~ Entrez2
    )
  ) %>%
  distinct(Protein1_entrez, Protein2_entrez, Score)

# Take the best-scoring edge per missing target
hippie_bf_best <- hippie_bf %>%
  group_by(Protein2_entrez) %>%
  slice_max(Score, n = 1, with_ties = FALSE) %>%
  ungroup()

cat("Missing targets recoverable via HIPPIE:", n_distinct(hippie_bf_best$Protein2_entrez), "\n")

# Resolve gene symbols
hippie_bf_best_chr <- hippie_bf_best %>%
  mutate(Protein1_entrez = as.character(Protein1_entrez),
         Protein2_entrez = as.character(Protein2_entrez))

hippie_bf_sym1 <- bitr(hippie_bf_best_chr$Protein1_entrez, fromType = "ENTREZID", toType = "SYMBOL",
                       OrgDb = org.Hs.eg.db, drop = TRUE) %>%
  dplyr::rename(Protein1_entrez = ENTREZID, Protein1 = SYMBOL) %>% distinct()
hippie_bf_sym2 <- bitr(hippie_bf_best_chr$Protein2_entrez, fromType = "ENTREZID", toType = "SYMBOL",
                       OrgDb = org.Hs.eg.db, drop = TRUE) %>%
  dplyr::rename(Protein2_entrez = ENTREZID, Protein2 = SYMBOL) %>% distinct()

hippie_bf_final <- hippie_bf_best_chr %>%
  left_join(hippie_bf_sym1, by = "Protein1_entrez") %>%
  left_join(hippie_bf_sym2, by = "Protein2_entrez") %>%
  dplyr::filter(!is.na(Protein1), !is.na(Protein2)) %>%
  transmute(
    Protein1, Protein2,
    Protein1_entrez,
    Protein2_entrez,
    Evidence = "HIPPIE protein-protein interaction (drug target expansion)",
    Source   = "HIPPIE_DrugTarget"
  ) %>%
  distinct()

cat("HIPPIE drug-target brute-force edges (with symbols):", nrow(hippie_bf_final), "\n")

still_unmappable <- missing_drug_targets %>%
  dplyr::filter(!target_entrez_id %in% hippie_bf_final$Protein2_entrez)
cat("Remaining unmappable targets:", nrow(still_unmappable), "\n")
```

# Step 5: Assemble Full Network v2 (Conservative)

```{r assemble_network_v2_conservative}
# Format KS pairs as network edges
ks_edges <- combined_with_omnipath %>%
  transmute(
    Protein1         = Kinase,
    Protein2         = Substrate,
    Protein1_entrez  = Kinase_entrez,
    Protein2_entrez  = Substrate_entrez,
    Evidence,
    Source
  ) %>%
  distinct()

# Combine all edge types
full_network_v2 <- bind_rows(
  ks_edges,
  hippie_ss_final,
  hippie_bf_final
) %>%
  distinct()

cat("=== Full Network v2 (Conservative) ===\n")
cat("Total edges:        ", nrow(full_network_v2), "\n")
cat("  KS pairs:         ", nrow(ks_edges), "\n")
cat("    Curated:        ", sum(ks_edges$Source %in% c("PhosphoSitePlus", "EPSD", "iPTMNet", "PhosphoNetworks", "PhosphoELM")), "\n")
cat("    Cantley (HPA):  ", sum(ks_edges$Source == "Cantley"), "\n")
cat("    OmniPath:       ", sum(ks_edges$Source == "OmniPath"), "\n")
cat("  HIPPIE SS:        ", nrow(hippie_ss_final), "\n")
cat("  HIPPIE DrugTarget:", nrow(hippie_bf_final), "\n")

all_nodes_v2 <- union(full_network_v2$Protein1_entrez, full_network_v2$Protein2_entrez)
cat("\nTotal unique nodes: ", length(all_nodes_v2), "\n")

# Summarize across sources
full_network_v2_summary <- full_network_v2 %>%
  group_by(Protein1, Protein2, Protein1_entrez, Protein2_entrez) %>%
  summarise(
    n_sources = n_distinct(Source),
    Sources   = paste(sort(unique(Source)), collapse = "; "),
    Evidence  = paste(sort(unique(Evidence)), collapse = "; "),
    .groups   = "drop"
  )

cat("\nSummary rows (unique protein pairs):", nrow(full_network_v2_summary), "\n")

# Save
write_csv(full_network_v2,         here("results/extended_kinome_network_v2.csv"))
write_csv(full_network_v2_summary, here("results/extended_kinome_network_summary_v2.csv"))
cat("\nSaved:\n")
cat("  results/extended_kinome_network_v2.csv\n")
cat("  results/extended_kinome_network_summary_v2.csv\n")
```

# Drug-Target Coverage Analysis

```{r coverage_analysis}
# Read v1 for comparison
network_v1 <- read_csv(here("results/extended_kinome_network_summary.csv"), show_col_types = FALSE)
nodes_v1 <- union(as.character(network_v1$Protein1_entrez),
                  as.character(network_v1$Protein2_entrez))

# Drug targets
drug_target_ids <- drug_targets$target_entrez_id
total <- length(drug_target_ids)

# Coverage
covered_v1 <- sum(drug_target_ids %in% nodes_v1)
covered_v2 <- sum(drug_target_ids %in% all_nodes_v2)

cat("=== Drug-Target Coverage ===\n")
cat("v1 coverage:", covered_v1, "/", total, sprintf("(%.1f%%)\n", 100*covered_v1/total))
cat("v2 coverage:", covered_v2, "/", total, sprintf("(%.1f%%)\n", 100*covered_v2/total))
cat("Improvement: +", covered_v2 - covered_v1, "targets (+",
    round(100*(covered_v2 - covered_v1)/total, 1), "ppts)\n", sep = "")
```

```{r coverage_viz}
# Layer-by-layer coverage
ks_nodes_v2 <- union(as.character(combined_with_omnipath$Kinase_entrez),
                     as.character(combined_with_omnipath$Substrate_entrez))
ss_nodes <- union(ks_nodes_v2, c(hippie_ss_final$Protein1_entrez, hippie_ss_final$Protein2_entrez))
bf_nodes <- union(ss_nodes, c(hippie_bf_final$Protein1_entrez, hippie_bf_final$Protein2_entrez))

layer_coverage <- tibble(
  layer = c("v1 (with HPA)", "v2: + OmniPath", "v2: + HIPPIE SS", "v2: + HIPPIE BF"),
  covered = c(
    covered_v1,
    sum(drug_target_ids %in% ks_nodes_v2),
    sum(drug_target_ids %in% ss_nodes),
    sum(drug_target_ids %in% bf_nodes)
  )
) %>%
  mutate(
    missing     = total - covered,
    coverage_pct = 100 * covered / total,
    layer = fct_inorder(layer)
  )

ggplot(layer_coverage, aes(x = layer)) +
  geom_col(aes(y = coverage_pct), fill = "#4C78A8", color = "grey20", width = 0.7) +
  geom_col(aes(y = 100 - coverage_pct), fill = "#E45756", color = "grey20",
           width = 0.7, position = position_stack(reverse = FALSE)) +
  geom_text(aes(y = coverage_pct / 2,
                label = paste0(covered, "\n(", round(coverage_pct, 1), "%)")),
            color = "white", fontface = "bold", size = 3.5) +
  geom_text(aes(y = coverage_pct + (100 - coverage_pct) / 2,
                label = paste0(missing, "\n(", round(100 - coverage_pct, 1), "%)")),
            fontface = "bold", size = 3.5) +
  scale_y_continuous(labels = function(x) paste0(x, "%"), limits = c(0, 101),
                     expand = expansion(mult = c(0, 0.01))) +
  labs(x = NULL, y = "% of drug targets",
       title = "Drug-target coverage: v2 conservative approach (HPA retained)") +
  theme_pubr(x.text.angle = 20) +
  theme(axis.text.x = element_text(size = 10))
```

## Network Statistics

```{r network_stats}
v1_stats <- tibble(
  metric = c("Total edges", "Unique kinases", "Unique substrates", "Total nodes"),
  v1 = c(
    nrow(network_v1),
    n_distinct(network_v1$Protein1),
    n_distinct(network_v1$Protein2),
    length(nodes_v1)
  ),
  v2 = c(
    nrow(full_network_v2_summary),
    n_distinct(full_network_v2_summary$Protein1),
    n_distinct(full_network_v2_summary$Protein2),
    length(all_nodes_v2)
  )
) %>%
  mutate(change = v2 - v1, pct_change = round(100 * change / v1, 1))

cat("=== Network Size: v1 vs v2 ===\n")
print(v1_stats)
```

## Source Composition

```{r source_composition}
edge_sources <- full_network_v2 %>%
  mutate(Source_clean = case_when(
    Source %in% c("PhosphoSitePlus", "EPSD", "iPTMNet", "PhosphoNetworks", "PhosphoELM") ~ "Curated",
    Source == "Cantley"      ~ "Cantley\n(with HPA)",
    Source == "OmniPath"     ~ "OmniPath",
    Source == "HIPPIE"       ~ "HIPPIE\n(substrate-substrate)",
    Source == "HIPPIE_DrugTarget" ~ "HIPPIE\n(drug target)",
    TRUE ~ Source
  )) %>%
  count(Source_clean) %>%
  mutate(pct = 100 * n / sum(n),
         Source_clean = fct_reorder(Source_clean, n))

ggplot(edge_sources, aes(x = Source_clean, y = n, fill = Source_clean)) +
  geom_col(color = "grey20", width = 0.7, show.legend = FALSE) +
  geom_text(aes(label = paste0(comma(n), "\n(", round(pct, 1), "%)")),
            vjust = -0.3, size = 3.2) +
  scale_fill_manual(values = c(
    "Curated"                = "#4C78A8",
    "Cantley\n(with HPA)"    = "#54A24B",
    "OmniPath"               = "#F58518",
    "HIPPIE\n(substrate-substrate)" = "#72B7B2",
    "HIPPIE\n(drug target)"  = "#B279A2"
  )) +
  scale_y_continuous(labels = comma, expand = expansion(mult = c(0, 0.15))) +
  labs(x = NULL, y = "Number of edges",
       title = "v2 network edge composition (conservative approach)") +
  theme_pubr(x.text.angle = 30) +
  theme(axis.text.x = element_text(size = 9))
```
