---
title: "Extended Kinome Network v4 — No HIPPIE"
subtitle: "Curated kinase-substrate + OmniPath signaling, without protein-protein interactions"
author: "Chinmaya Joisa"
date: today
format:
  html:
    page-layout: full
    embed-resources: true
    code-fold: true
execute:
  echo: true
  cache: false
  message: false
  warning: false
---

```{r setup, include=FALSE}
require("knitr")
knitr::opts_knit$set(root.dir = here::here())
set.seed(123)
```

```{r libraries}
library(tidyverse)
library(fst)
library(here)
library(data.table)
library(igraph)
invisible(gc())
```

# v4 Extended Kinome Network: Curated KS Only (No HIPPIE)

**Network architecture (v4):**
- Layer 1: Curated kinase-substrate pairs (PSP + EPSD + iPTMNet + PhosphoELM + PhosphoNetworks, experimental)
- Layer 2: Cantley predicted kinase-substrate (with HPA co-localization filtering)
- Layer 3: OmniPath enzyme-substrate (PhosphoSite + SIGNOR curated, S/T/Y only)
- ~~Layer 4: HIPPIE substrate-substrate PPIs~~ **REMOVED**
- Layer 4: OmniPath directed signaling (for non-KS drug targets)
- ~~Layer 5: HIPPIE 1-hop drug-target expansion~~ **REMOVED**

**Key differences from v3:**
- v3 uses HIPPIE PPIs (68.5% of edges) to connect substrates and to bridge non-kinase drug targets
- v4 removes all HIPPIE, keeping only **curated kinase-substrate** and **curated signaling**
- v4 is smaller and more conservative, with reduced drug-target coverage
- v4 focus: mechanistically grounded interactions only (phosphorylation + directed signaling)


## Load Data

```{r load_data}
# Drug targets
drug_targets <- read.fst(here("data/kinome_profiling_novartis_evebio_combined.fst")) %>%
  as_tibble() %>%
  distinct(target, target_entrez_id) %>%
  mutate(target_entrez_id = as.character(target_entrez_id))

# Load v3 network (to get backbone structure)
v3_summary <- read_csv(here("results/extended_kinome_network_summary_v3.csv"),
                       show_col_types = FALSE) %>%
  dplyr::filter(!is.na(Protein1), !is.na(Protein2)) %>%
  mutate(
    Protein1_entrez = as.character(Protein1_entrez),
    Protein2_entrez = as.character(Protein2_entrez)
  )

v3_full <- read_csv(here("results/extended_kinome_network_v3.csv"),
                    show_col_types = FALSE) %>%
  dplyr::filter(!is.na(Protein1), !is.na(Protein2)) %>%
  mutate(
    Protein1_entrez = as.character(Protein1_entrez),
    Protein2_entrez = as.character(Protein2_entrez)
  )

cat("Data loaded.\n")
cat("  Drug targets:  ", nrow(drug_targets), "\n")
cat("  v3 full edges: ", nrow(v3_full), "\n")
cat("  v3 pairs:      ", nrow(v3_summary), "\n")
```

## Build v4: Remove HIPPIE Layers from v3

```{r build_v4}
# v4 = v3 without HIPPIE edges (both general PPIs and drug-target 1-hop)
v4_full <- v3_full %>%
  dplyr::filter(Source != "HIPPIE", Source != "HIPPIE_DrugTarget")

v4_summary <- v3_summary %>%
  dplyr::filter(Sources != "HIPPIE", !grepl("HIPPIE", Sources))

# Recompute summary (deduplicated pairs with aggregated sources)
v4_summary_recomputed <- v4_full %>%
  group_by(Protein1, Protein2, Protein1_entrez, Protein2_entrez) %>%
  summarise(
    Sources = paste(unique(Source), collapse = "; "),
    Evidence = NA_character_,
    n_sources = n(),
    .groups = "drop"
  )

cat("=== v4 Network (No HIPPIE) ===\n")
cat("Full edges:        ", nrow(v4_full), "\n")
cat("Unique pairs:      ", nrow(v4_summary_recomputed), "\n")

# Nodes and coverage
nodes_v4 <- union(v4_summary_recomputed$Protein1_entrez, 
                  v4_summary_recomputed$Protein2_entrez)
cat("Total nodes:       ", length(nodes_v4), "\n")

drug_targets_in_v4 <- sum(drug_targets$target_entrez_id %in% nodes_v4)
pct_v4 <- 100 * drug_targets_in_v4 / nrow(drug_targets)
cat("Drug targets:      ", drug_targets_in_v4, "/", nrow(drug_targets), 
    sprintf("(%.1f%%)\n", pct_v4))
```

## Export v4 Results

```{r export_v4}
write_csv(v4_full, here("results/extended_kinome_network_v4.csv"))
write_csv(v4_summary_recomputed, here("results/extended_kinome_network_summary_v4.csv"))

cat("\nExported:\n")
cat("  results/extended_kinome_network_v4.csv (", nrow(v4_full), " edges)\n")
cat("  results/extended_kinome_network_summary_v4.csv (", nrow(v4_summary_recomputed), " pairs)\n")
```

## v3 vs v4 Comparison

```{r comparison}
v3_nodes <- union(v3_summary$Protein1_entrez, v3_summary$Protein2_entrez)
v3_drug_targets <- sum(drug_targets$target_entrez_id %in% v3_nodes)

cat("\n=== v3 vs v4 Comparison ===\n\n")

cat("Network size:\n")
cat("  v3: ", nrow(v3_summary), " edges, ", length(v3_nodes), " nodes\n")
cat("  v4: ", nrow(v4_summary_recomputed), " edges, ", length(nodes_v4), " nodes\n")
cat("  Difference: ", nrow(v3_summary) - nrow(v4_summary_recomputed), " edges (HIPPIE removed)\n\n")

cat("Drug-target coverage:\n")
cat("  v3: ", v3_drug_targets, "/", nrow(drug_targets), 
    sprintf(" (%.1f%%)\n", 100 * v3_drug_targets / nrow(drug_targets)))
cat("  v4: ", drug_targets_in_v4, "/", nrow(drug_targets), 
    sprintf(" (%.1f%%)\n", pct_v4))
cat("  Coverage loss: ", v3_drug_targets - drug_targets_in_v4, " targets\n\n")

# Show which drug targets are missing in v4 (only in HIPPIE)
missing_in_v4_drug_targets <- drug_targets %>%
  dplyr::filter(target_entrez_id %in% v3_nodes,
                !target_entrez_id %in% nodes_v4) %>%
  pull(target)

cat("Drug targets lost in v4 (only connected via HIPPIE):\n")
cat("  Count: ", length(missing_in_v4_drug_targets), "\n")
if (length(missing_in_v4_drug_targets) > 0) {
  cat("  Examples:", paste(head(missing_in_v4_drug_targets, 10), collapse = ", "), "\n")
}
```
