---
title: "Extended Kinome Network v4 — Network EDA"
subtitle: "Drug target importance, hub analysis, and community structure"
author: "Chinmaya Joisa"
date: today
toc: true
toc-depth: 3
toc-title: Table of Contents
format:
  html:
    page-layout: full
    embed-resources: true
    code-fold: true
    code-tools: true
execute:
  echo: true
  cache: false
  message: false
  warning: false
  fig-align: center
  fig-dpi: 300
---

```{r setup, include=FALSE}
require("knitr")
knitr::opts_knit$set(root.dir = here::here())
set.seed(123)
```

```{r libraries}
library(tidyverse)
library(fst)
library(here)
library(scales)
library(ggpubr)
library(patchwork)
library(igraph)
library(DT)
invisible(gc())
```

# Overview

Network exploratory data analysis (EDA) of the v4 extended kinome network (curated kinase-substrate only, no HIPPIE). Focus areas:
- **Drug-target importance**: Which drug targets are most influential by centrality measures?
- **Non-drug target hubs**: Which backbone proteins are critical bridges?
- **Network communities**: Leiden clustering to identify functional modules
- **Drug target distribution**: How drug targets distribute across communities
- **Source quality**: Breakdown of edges by curation source

# 1. Load Data

```{r load_v4}
# Load v4 network
network_v4 <- read_csv(here("results/extended_kinome_network_summary_v4.csv"),
                       show_col_types = FALSE) %>%
  dplyr::filter(!is.na(Protein1), !is.na(Protein2),
                Protein1 != "", Protein2 != "") %>%
  mutate(Protein1_entrez = as.character(Protein1_entrez),
         Protein2_entrez = as.character(Protein2_entrez))

network_v4_full <- read_csv(here("results/extended_kinome_network_v4.csv"),
                            show_col_types = FALSE) %>%
  dplyr::filter(!is.na(Protein1), !is.na(Protein2),
                Protein1 != "", Protein2 != "") %>%
  mutate(Protein1_entrez = as.character(Protein1_entrez),
         Protein2_entrez = as.character(Protein2_entrez))

# Drug target panel
drug_targets <- read.fst(here("data/kinome_profiling_novartis_evebio_combined.fst")) %>%
  as_tibble() %>%
  distinct(target, target_entrez_id) %>%
  mutate(target_entrez_id = as.character(target_entrez_id),
         is_drug_target = TRUE)

# Identify drug targets in network
drug_target_set <- drug_targets$target_entrez_id

nodes_v4 <- union(network_v4$Protein1_entrez, network_v4$Protein2_entrez)
drug_targets_in_v4 <- drug_targets %>%
  dplyr::filter(target_entrez_id %in% nodes_v4)

cat("v4 Network overview:\n")
cat("  Edges:               ", nrow(network_v4), "\n")
cat("  Nodes:               ", length(nodes_v4), "\n")
cat("  Drug targets:        ", nrow(drug_targets), "\n")
cat("  Drug targets in v4:  ", nrow(drug_targets_in_v4), 
    sprintf(" (%.1f%%)\n", 100 * nrow(drug_targets_in_v4) / nrow(drug_targets)))
cat("  Non-drug nodes:      ", length(nodes_v4) - nrow(drug_targets_in_v4), "\n")
```

# 2. Build Network Graph

```{r build_graph}
build_igraph <- function(net) {
  graph_from_data_frame(
    net %>% dplyr::select(from = Protein1, to = Protein2),
    directed = FALSE
  )
}

g_v4 <- build_igraph(network_v4)

# Basic stats
comps <- igraph::components(g_v4)
comp_sizes <- sort(comps$csize, decreasing = TRUE)
degs <- igraph::degree(g_v4)

cat("=== v4 Network-Wide Properties ===\n")
cat("Nodes:                    ", vcount(g_v4), "\n")
cat("Edges:                    ", ecount(g_v4), "\n")
cat("Density:                  ", round(edge_density(g_v4), 5), "\n")
cat("Number of components:     ", comps$no, "\n")
cat("Largest component nodes:  ", comp_sizes[1], " (", 
    round(100 * comp_sizes[1] / vcount(g_v4), 1), "%)\n")
cat("Global clustering coeff:  ", round(igraph::transitivity(g_v4, type = "global"), 4), "\n\n")

cat("Degree statistics:\n")
cat("  Mean:   ", round(mean(degs), 1), "\n")
cat("  Median: ", median(degs), "\n")
cat("  Max:    ", max(degs), "\n")
```

# 3. Drug-Target Importance: Centrality Measures

```{r drug_target_centrality}
# Create mapping from gene symbol to Entrez ID
symbol_to_entrez <- bind_rows(
  network_v4 %>% dplyr::select(symbol = Protein1, entrez = Protein1_entrez),
  network_v4 %>% dplyr::select(symbol = Protein2, entrez = Protein2_entrez)
) %>%
  distinct(symbol, entrez)

# Compute centrality measures
betweenness <- igraph::betweenness(g_v4, normalized = TRUE)
closeness <- igraph::closeness(g_v4, normalized = TRUE)
eigenvector <- igraph::eigen_centrality(g_v4)$vector
pagerank <- igraph::page_rank(g_v4)$vector

centrality_df <- tibble(
  node = names(degs),
  degree = as.numeric(degs),
  betweenness = betweenness[names(degs)],
  closeness = closeness[names(degs)],
  eigenvector = eigenvector[names(degs)],
  pagerank = pagerank[names(degs)]
) %>%
  left_join(symbol_to_entrez, by = c("node" = "symbol")) %>%
  mutate(
    is_drug_target = entrez %in% drug_target_set,
    degree_rank = rank(-degree),
    betweenness_rank = rank(-betweenness),
    closeness_rank = rank(-closeness),
    eigenvector_rank = rank(-eigenvector),
    pagerank_rank = rank(-pagerank),
    combined_rank = degree_rank + betweenness_rank + closeness_rank + 
                    eigenvector_rank + pagerank_rank
  )

# Top drug targets by importance
top_drug_targets <- centrality_df %>%
  dplyr::filter(is_drug_target) %>%
  slice_min(combined_rank, n = 15)

cat("=== Top 15 Drug Targets by Importance (combined rank) ===\n")
top_drug_targets %>%
  dplyr::select(node, degree, betweenness, closeness, eigenvector, pagerank, combined_rank) %>%
  arrange(combined_rank) %>%
  print()

# Centrality comparison: drug targets vs non-drug nodes
comparison_stats <- centrality_df %>%
  group_by(is_drug_target) %>%
  summarise(
    n_nodes = n(),
    mean_degree = mean(degree),
    mean_betweenness = mean(betweenness),
    mean_closeness = mean(closeness),
    mean_eigenvector = mean(eigenvector),
    mean_pagerank = mean(pagerank),
    .groups = "drop"
  ) %>%
  mutate(node_type = if_else(is_drug_target, "Drug targets", "Backbone nodes"))

cat("\nCentrality: Drug targets vs Backbone nodes\n")
print(comparison_stats)
```

```{r plot_centrality_dist}
cent_long <- centrality_df %>%
  pivot_longer(cols = c(degree, betweenness, closeness, eigenvector, pagerank),
               names_to = "metric", values_to = "value") %>%
  mutate(
    metric = factor(metric, levels = c("degree", "betweenness", "closeness", "eigenvector", "pagerank")),
    node_type = if_else(is_drug_target, "Drug target", "Backbone")
  )

ggplot(cent_long, aes(x = node_type, y = value, fill = node_type)) +
  geom_boxplot(color = "grey20", width = 0.5, show.legend = FALSE, alpha = 0.8) +
  facet_wrap(~metric, scales = "free_y") +
  scale_fill_manual(values = c("Drug target" = "#E15759", "Backbone" = "#4C78A8")) +
  labs(x = NULL, y = "Centrality value", 
       title = "Centrality measures: drug targets vs backbone nodes") +
  theme_pubr() +
  theme(axis.text.x = element_text(angle = 15, hjust = 1))
```

# 4. Non-Drug Target Hubs

```{r non_drug_hubs}
backbone_hubs <- centrality_df %>%
  dplyr::filter(!is_drug_target) %>%
  slice_min(combined_rank, n = 20)

cat("=== Top 20 Backbone Hubs (non-drug targets) ===\n")
backbone_hubs %>%
  dplyr::select(node, degree, betweenness, closeness, eigenvector, pagerank, combined_rank) %>%
  arrange(combined_rank) %>%
  print()

# Plot top backbone hubs
ggplot(backbone_hubs, aes(x = reorder(node, -combined_rank), y = degree)) +
  geom_col(fill = "#4C78A8", color = "grey20") +
  geom_text(aes(label = degree), vjust = -0.3, size = 3) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
  coord_flip() +
  labs(x = NULL, y = "Degree",
       title = "Top 20 backbone hub proteins (non-drug targets)") +
  theme_pubr()
```

# 5. Louvain Community Detection

```{r leiden_clustering}
cat("Running community detection (Louvain method)...\n")

# Use largest component for better clustering
largest_comp_nodes <- which(comps$membership == which.max(comps$csize))
g_main <- igraph::induced_subgraph(g_v4, largest_comp_nodes)

# Run Louvain community detection (igraph implementation)
communities <- igraph::cluster_louvain(g_main)

cat("Community detection complete.\n")
cat("  Number of communities: ", length(communities), "\n")
cat("  Modularity:            ", round(modularity(communities), 4), "\n")
cat("  Main component size:   ", vcount(g_main), "\n")

# Add cluster assignments to centrality data
cluster_map <- tibble(
  node = names(V(g_main)),
  cluster = as.integer(membership(communities))
)

centrality_clustered <- centrality_df %>%
  left_join(cluster_map, by = "node") %>%
  replace_na(list(cluster = 0))  # 0 for nodes not in largest component

cat("\nCluster sizes:\n")
cluster_sizes <- centrality_clustered %>%
  count(cluster, name = "n_nodes") %>%
  mutate(n_drug_targets = NA_integer_,
         pct_drug_targets = NA_real_)

for (i in cluster_sizes$cluster) {
  n_dt <- sum(centrality_clustered$is_drug_target & centrality_clustered$cluster == i)
  n_tot <- sum(centrality_clustered$cluster == i)
  cluster_sizes$n_drug_targets[cluster_sizes$cluster == i] <- n_dt
  if (n_tot > 0) {
    cluster_sizes$pct_drug_targets[cluster_sizes$cluster == i] <- 100 * n_dt / n_tot
  }
}

print(cluster_sizes %>% arrange(desc(n_nodes)) %>% head(15))
```

```{r plot_cluster_composition}
# Drug target distribution across clusters
cluster_composition <- centrality_clustered %>%
  dplyr::filter(cluster > 0) %>%  # Exclude isolated nodes
  group_by(cluster) %>%
  summarise(
    n_drug = sum(is_drug_target),
    n_backbone = sum(!is_drug_target),
    n_total = n(),
    pct_drug = 100 * n_drug / n_total,
    .groups = "drop"
  ) %>%
  slice_max(n_total, n = 15) %>%
  arrange(desc(n_total))

ggplot(cluster_composition, aes(x = reorder(factor(cluster), -n_total), y = n_total, 
                                fill = pct_drug)) +
  geom_col(color = "grey20") +
  geom_text(aes(label = paste0(n_drug, " drug\n", n_backbone, " backbone")),
            vjust = 0.5, hjust = 0.5, size = 2.5, color = "white") +
  scale_fill_gradient(low = "#4C78A8", high = "#E15759", name = "% Drug targets") +
  labs(x = "Community", y = "Node count",
       title = "Louvain communities: size and drug-target composition") +
  theme_pubr()
```

# 6. Drug Targets Across Clusters

```{r drug_targets_per_cluster}
drug_targets_in_clusters <- centrality_clustered %>%
  dplyr::filter(is_drug_target, cluster > 0) %>%
  arrange(desc(degree))

cat("Drug targets per community:\n")
drug_targets_in_clusters %>%
  count(cluster, name = "n_drug_targets") %>%
  arrange(desc(n_drug_targets)) %>%
  print()

# Identify which drug targets are in which clusters
dt_cluster_map <- drug_targets_in_clusters %>%
  left_join(drug_targets, by = c("node" = "target_entrez_id")) %>%
  dplyr::select(gene = target, cluster, degree, betweenness, eigenvector) %>%
  arrange(cluster, desc(degree))

cat("\nDrug target community assignments (sorted by degree):\n")
dt_cluster_map %>%
  group_by(cluster) %>%
  slice_max(degree, n = 5) %>%
  print()
```

# 7. Source Quality Breakdown

```{r source_breakdown}
source_quality <- tibble(
  source_name = c("Curated_KS", "Cantley", "OmniPath", "OmniPath_Signaling"),
  quality_tier = c("Curated KS", "Predicted (HPA-validated)", "Curated enzyme-substrate",
                   "Curated directed signaling")
)

# Parse sources from the Sources column
edge_sources_parsed <- network_v4_full %>%
  separate_rows(Source, sep = "; ") %>%
  mutate(Source = str_trim(Source)) %>%
  count(Source, name = "n_edges") %>%
  left_join(source_quality, by = c("Source" = "source_name")) %>%
  mutate(
    quality_tier = replace_na(quality_tier, "Other"),
    pct = round(100 * n_edges / sum(n_edges), 1),
    Source = fct_reorder(Source, n_edges)
  )

ggplot(edge_sources_parsed, aes(x = Source, y = n_edges, fill = quality_tier)) +
  geom_col(color = "grey20", width = 0.7) +
  geom_text(aes(label = paste0(comma(n_edges), "\n(", pct, "%)")),
            hjust = -0.05, size = 3) +
  scale_fill_manual(values = c(
    "Curated KS"                    = "#4C78A8",
    "Predicted (HPA-validated)"     = "#54A24B",
    "Curated enzyme-substrate"      = "#F58518",
    "Curated directed signaling"    = "#B279A2"
  )) +
  scale_y_continuous(labels = comma, expand = expansion(mult = c(0, 0.25))) +
  coord_flip() +
  labs(x = NULL, y = "Number of edges",
       title = "v4 network edges by source and quality tier",
       fill = "Quality tier") +
  theme_pubr() +
  theme(legend.position = "bottom", legend.direction = "vertical")
```

# 8. Degree Distribution

```{r degree_dist}
deg_df <- tibble(
  degree = as.numeric(degs),
  is_drug_target = names(degs) %in% drug_target_set
)

ggplot(deg_df, aes(x = degree + 1, fill = is_drug_target)) +
  geom_histogram(bins = 30, color = "grey20", alpha = 0.8) +
  scale_x_log10(labels = label_comma(), breaks = scales::trans_breaks("log10", function(x) 10^x)) +
  scale_fill_manual(values = c("TRUE" = "#E15759", "FALSE" = "#4C78A8"),
                    labels = c("TRUE" = "Drug target", "FALSE" = "Backbone")) +
  labs(x = "Node degree + 1 (log scale)", y = "Count",
       title = "v4 network degree distribution",
       fill = NULL) +
  theme_pubr() +
  theme(legend.position = "bottom")
```

# 9. Summary Report

```{r summary}
cat("======================================================================\n")
cat("EXTENDED KINOME NETWORK v4 — EDA SUMMARY\n")
cat("======================================================================\n\n")

cat("A. Network overview\n")
cat("   Edges:  ", nrow(network_v4), "\n")
cat("   Nodes:  ", length(nodes_v4), "\n")
cat("   Drug targets:  ", nrow(drug_targets_in_v4), "/", nrow(drug_targets),
    sprintf(" (%.1f%%)\n", 100 * nrow(drug_targets_in_v4) / nrow(drug_targets)))
cat("   Mean degree:   ", round(mean(degs), 1), "\n\n")

cat("B. Architecture (curated only, no HIPPIE PPIs)\n")
cat("   Kinase-substrate (curated):    experimental evidence from PSP/EPSD/iPTMNet/etc.\n")
cat("   Kinase-substrate (predicted):  Cantley atlas with HPA co-localization\n")
cat("   Enzyme-substrate (curated):    OmniPath/PhosphoSite/SIGNOR\n")
cat("   Directed signaling:            OmniPath (SIGNOR/HPRD/SPIKE) for drug targets\n\n")

cat("C. Drug-target centrality\n")
cat("   Mean degree (drug):     ", 
    round(mean(centrality_df$degree[centrality_df$is_drug_target]), 2), "\n")
cat("   Mean degree (backbone): ", 
    round(mean(centrality_df$degree[!centrality_df$is_drug_target]), 2), "\n")
cat("   Mean betweenness (drug):     ", 
    round(mean(centrality_df$betweenness[centrality_df$is_drug_target]), 4), "\n")
cat("   Mean betweenness (backbone): ", 
    round(mean(centrality_df$betweenness[!centrality_df$is_drug_target]), 4), "\n\n")

cat("D. Louvain community structure\n")
n_clusters <- length(unique(centrality_clustered$cluster[centrality_clustered$cluster > 0]))
cat("   Number of communities: ", n_clusters, "\n")
cat("   Largest community:     ", max(cluster_sizes$n_nodes), " nodes\n\n")

cat("======================================================================\n")
cat("KEY INSIGHTS\n")
cat("======================================================================\n")
cat("1. Drug-target importance varies widely across centrality metrics\n")
cat("2. Backbone hubs are critical for signal propagation through the network\n")
cat("3. Louvain community detection reveals functional modules with varying drug-target density\n")
cat("4. v4 (curated-only) is more conservative than v3 but mechanistically grounded\n")
cat("======================================================================\n")
```
