---
title: "Analysis and Integration of The Specificity Atlases and Curated Kinase-Substrate Data"
editor: visual
author: "Chinmaya Joisa"
date: "08/15/2025"
toc: true
toc-depth: 5
toc-title: Table of Contents
highlight-style: pygments
format: 
  html:
    embed-resources: true
    code-fold: true
    code-tools: true
execute:
  echo: false
  cache: true
  message: false
  warning: false
  out-width: "100%"
  fig-align: center
  fig-dpi: 300
---

```{r setup, include=FALSE, cache = FALSE}
require("knitr")
## setting working directory
knitr::opts_knit$set(root.dir = here::here())
```

```{r, include=FALSE}

library(tidyverse)
library(here)
library(readxl)
library(DarkKinaseTools)
library(biomaRt)
library(clipr)
library(conflicted)
library(tidymodels)
library(pROC)
library(glue)
library(clusterProfiler)
library(org.Hs.eg.db)
library(AnnotationDbi)
library(data.table)
library(fastmatch) 
library(Biostrings)
library(httr2)
library(tidytext)

library(tidygraph)
library(ggraph)
library(igraph)
library(ggpubr)
library(ggrepel)
library(scales)
library(patchwork)
library(cowplot)
library(ggupset)
library(ggalluvial)
library(ggtext)
library(magick)

all_kinases = all_kinases

pal_src <- c(
  "PhosphoSitePlus" = "#4C78A8",
  "PhosphoELM"      = "#F58518",
  "PhosphoNetworks" = "#54A24B",
  "Specificity Atlas"= "#B279A2",
  "Specificity Atlas≥99%" = "#B279A2"
)

pal_role <- c(Kinase="#2F4B7C", Substrate="#F95D6A", `Kinase,Substrate`="#2CA02C")

theme_pub <- function(base_size = 11){
  theme_pubr(base_size = base_size) %+replace%
    theme(
      panel.grid.major = element_line(color="grey90", linewidth=0.2),
      panel.grid.minor = element_blank(),
      axis.title = element_text(face="bold"),
      plot.title = element_text(face="bold", hjust=0),
      legend.title = element_text(face="bold")
    )
}


conflict_prefer("select", "dplyr")
conflict_prefer("filter", "dplyr")
conflict_prefer("rename", "dplyr")
conflicts_prefer(base::setdiff)
conflicts_prefer(base::union)
conflicts_prefer(dplyr::coalesce)
conflicts_prefer(dplyr::slice)
conflicts_prefer(dplyr::first)
conflicts_prefer(base::as.factor)
conflicts_prefer(base::intersect)
conflicts_prefer(stringr::fixed)
conflicts_prefer(dplyr::desc)
```
    
```{r} 
#Read in data
    # Read in Cantley Atlas data
    cantley_atlas <- read_csv(here("data/CantleyData/CantleyAtlas.csv"))
    
    cantley_tyrosine = read_csv(here("data/CantleyData/CantleyTyrosine.csv"))

    # Read in PhosphoSitePlus data
    # phosphosite_plus <- read_tsv(here("data/Kinase_Substrate_Dataset"), skip = 3)

    # phospho_elm = read_tsv(here("data/phosphoELM_all_2015-04.dump")) %>%
    #   mutate(
    #     kinases = case_when(
    #       kinases == "Lck" ~ "LCK",
    #       kinases == "Fyn" ~ "FYN",
    #       kinases == "Abl" ~ "ABL1",
    #       kinases == "Aurora B" ~ "AURKB",
    #       kinases == "PDK-1" ~ "PDPK1",
    #       kinases == "PKC_theta" ~ "PRKCQ",
    #       kinases == "IKK_beta" ~ "IKBKB",
    #       kinases == "CaM-KII_alpha" ~ "CAMK2A",
    #       kinases == "PKC_alpha" ~ "PRKCA",
    #       kinases == "CK2_alpha" ~ "CSNK2A1",
    #       kinases == "GSK-3_alpha" ~ "GSK3A",
    #       kinases == "GSK-3_beta" ~ "GSK3B",
    #       kinases == "PKB_beta" ~ "AKT2",
    #       kinases == "DNA-PK" ~ "PRKDC",
    #       kinases == "ErbB2" ~ "ERBB2",
    #       kinases == "Aurora A" ~ "AURKA",
    #       kinases == "PKC_delta" ~ "PRKCD",
    #       kinases == "Csk" ~ "CSK",
    #       kinases == "PDGFR_beta" ~ "PDGFRB",
    #       kinases == "Mnk1" ~ "MKNK1",
    #       kinases == "Mnk2" ~ "MKNK2",
    #       kinases == "GRK-2" ~ "ADRBK1",
    #       kinases == "GRK-5" ~ "GRK5",
    #       kinases == "CK1_alpha" ~ "CSNK1A1",
    #       kinases == "Lyn" ~ "LYN",
    #       kinases == "PDHK1" ~ "PDK1",
    #       kinases == "PDHK3" ~ "PDK3",
    #       kinases == "PDHK4" ~ "PDK4",
    #       kinases == "PKC_epsilon" ~ "PRKCE",
    #       kinases == "PKG1/cGK-I" ~ "PRKG1",
    #       kinases == "Fgr" ~ "FGR",
    #       kinases == "p70S6K" ~ "RPS6KB1",
    #       kinases == "PKC_beta" ~ "PRKCB",
    #       kinases == "CK1_delta" ~ "CSNK1D",
    #       kinases == "CK2_beta" ~ "CSNK2B",
    #       kinases == "RSK-3" ~ "RPS6KA2",
    #       kinases == "RSK-2" ~ "RPS6KA3",
    #       kinases == "CaM-KIV" ~ "CAMK4",
    #       kinases == "Tyk2" ~ "TYK2",
    #       kinases == "RSK-1" ~ "RPS6KA1",
    #       kinases == "IKK_alpha" ~ "CHUK",
    #       kinases == "mTOR" ~ "MTOR",
    #       kinases == "CK1_epsilon" ~ "CSNK1E",
    #       kinases == "PIM-1" ~ "PIM1",
    #       kinases == "Eg3 kinase" ~ "MELK",
    #       kinases == "GRK-4" ~ "GRK4",
    #       kinases == "PKC_zeta" ~ "PRKCZ",
    #       kinases == "PDKC" ~ "PRKCH",
    #       kinases == "Chak1" ~ "BRSK1",
    #       kinases == "TGFbR2" ~ "TGFBR2",
    #       kinases == "Brk" ~ "PTK6",
    #       kinases == "RSK-5" ~ "RPS6KA5",
    #       kinases == "IKK_epsilon" ~ "IKBKE",
    #       kinases == "TGFbR1" ~ "TGFBR1",
    #       kinases == "Fes" ~ "FES",
    #       kinases == "PKA_alpha" ~ "PRKACA",
    #       kinases == "EphA3" ~ "EPHA3",
    #       kinases == "EphA4" ~ "EPHA4",
    #       kinases == "EphA8" ~ "EPHA8",
    #       kinases == "Kit" ~ "KIT",
    #       kinases == "EphB3" ~ "EPHB3",
    #       kinases == "Mer" ~ "MERTK",
    #       kinases == "PKC_eta" ~ "PRKCH",
    #       kinases == "PDK-2" ~ "PDPK1",  # ambiguous, often a site not a gene
    #       kinases == "GRK-6" ~ "GRK6",
    #       kinases == "Titin kinase" ~ "TTN",
    #       kinases == "CaM-KI_alpha" ~ "CAMK1",
    #       kinases == "GRK-1" ~ "GRK1",
    #       kinases == "EphB2" ~ "EPHB2",
    #       kinases == "PDGFR_alpha" ~ "PDGFRA",
    #       kinases == "CCDPK" ~ "CDK5",
    #       kinases == "Axl" ~ "AXL",
    #       kinases == "GRK-3" ~ "ADRBK2",
    #       kinases == "MRCKa" ~ "CDC42BPA",
    #       kinases == "PKG2/cGK-II" ~ "PRKG2",
    #       kinases == "NuaK1" ~ "NUAK1",
    #       kinases == "Abl2" ~ "ABL2",
    #       kinases == "PKC_gamma" ~ "PRKCG",
    #       kinases == "PKC_iota" ~ "PRKCI",
    #       kinases == "CaM-KK_alpha" ~ "CAMKK1",
    #       kinases == "EphA2" ~ "EPHA2",
    #       TRUE ~ kinases # fallback to original if no match
    #     )
    #   )

    subcellular_localization = read_tsv(here("data/CantleyData/subcellular_location.tsv"))

    #phospho_networks <- read_csv(here("data/phospho_networks_kinase_substrate_sites.csv"))

    kinet <- read_csv(here("data/kinet_data/ksi_source_full_dataset.csv"))

    ochoa_fs = read_excel(here("data/41587_2019_344_MOESM5_ESM.xlsx"), sheet = "functional_score")

    ochoa = read_excel(here("data/41587_2019_344_MOESM4_ESM.xlsx"), sheet = "annotated_phosphoproteome")

    kinase_list_ref = read_excel(here("data/Comprehensive_Kinase_Simple.xlsx")) %>%
      select(Kinase = Gene, Kinase_Uniprot_ID = `Human Uniprot`, Family = `Kinase Family`,
             ref_entrez = `Human Gene ID`) %>%
      distinct()

    # The kinome table uses some superseded symbols (ICK, PRPF4B, COL4A3BP, STK19),
    # whereas the pipeline reports current HGNC symbols (CILK1, PRP4K, CERT1, WHR1).
    # Add a row under the current symbol as well, so family lookups by symbol keep
    # working for renamed genes.
    kinase_current_symbol <- AnnotationDbi::mapIds(
      org.Hs.eg.db, keys = as.character(kinase_list_ref$ref_entrez), column = "SYMBOL",
      keytype = "ENTREZID", multiVals = "first")
    kinase_list <- kinase_list_ref %>%
      mutate(current_symbol = unname(kinase_current_symbol[as.character(ref_entrez)])) %>%
      { bind_rows(select(., Kinase, Kinase_Uniprot_ID, Family),
                  filter(., !is.na(current_symbol), current_symbol != Kinase) %>%
                    transmute(Kinase = current_symbol, Kinase_Uniprot_ID, Family)) } %>%
      distinct(Kinase, .keep_all = TRUE)

```    

### Protein name matching
```{r}

# Helper: normalize UniProt accessions (drop isoform suffix)
norm_uniprot <- function(x) str_replace(x, "-\\d+$", "")

# Build canonical symbol map for a vector of gene-like names.
#
# NOTE: the previous version of this helper fell back to an ALIAS -> ENTREZ
# lookup and kept every hit. Because many legacy protein names are aliases of
# several genes, that one-to-many join silently duplicated a protein's data onto
# unrelated genes. For example "SMAP" is an alias of C11orf58 (UniProt O00193),
# of KIFAP3 ("Smg GDS-associated protein") and of BRD8, so a single O00193
# phosphosite was emitted three times under three different substrate names
# while keeping O00193's residue numbering. We now accept an ALIAS only when it
# resolves to exactly one gene, and let callers resolve the remainder from the
# UniProt accession (see resolve_by_uniprot()).
build_symbol_map <- function(x_symbols) {
  x_symbols <- unique(na.omit(x_symbols))

  # 1) direct SYMBOL -> ENTREZ
  m1 <- suppressWarnings(tryCatch(
    bitr(x_symbols, fromType = "SYMBOL", toType = "ENTREZID", OrgDb = org.Hs.eg.db),
    error = function(e) NULL))
  if (is.null(m1)) m1 <- tibble::tibble(SYMBOL = character(), ENTREZID = character())
  # a current symbol should be unique; drop it if it is not
  m1 <- m1 %>% add_count(SYMBOL) %>% filter(n == 1) %>% select(-n)

  # 2) remaining via ALIAS -> ENTREZ, UNAMBIGUOUS MATCHES ONLY
  left <- setdiff(x_symbols, m1$SYMBOL)
  m2 <- if (length(left)) suppressWarnings(tryCatch(
    bitr(left, fromType = "ALIAS", toType = "ENTREZID", OrgDb = org.Hs.eg.db),
    error = function(e) NULL)) else NULL
  if (is.null(m2)) {
    m2 <- tibble::tibble(SYMBOL = character(), ENTREZID = character())
  } else {
    m2 <- m2 %>%
      rename(SYMBOL = ALIAS) %>%
      distinct(SYMBOL, ENTREZID) %>%
      add_count(SYMBOL) %>%
      filter(n == 1) %>%          # <- the fix: reject ambiguous aliases
      select(-n)
  }

  mm <- bind_rows(m1, m2) %>% distinct()

  # canonical SYMBOL for each ENTREZ
  can_sym <- suppressMessages(AnnotationDbi::select(
    org.Hs.eg.db, keys = unique(mm$ENTREZID), keytype = "ENTREZID", columns = "SYMBOL"))
  left_join(mm, can_sym, by = "ENTREZID") %>%
    transmute(input_symbol = SYMBOL.x, ENTREZID, SYMBOL = SYMBOL.y) %>%
    distinct()
}

# Resolve gene identity from UniProt accessions, keeping only accessions that
# map to exactly one gene. Used as the second line of defence for substrates
# whose atlas gene label is an obsolete or ambiguous alias.
resolve_by_uniprot <- function(accessions) {
  accessions <- unique(na.omit(norm_uniprot(accessions)))
  if (!length(accessions)) {
    return(tibble::tibble(Substrate_Uniprot_ID = character(),
                          ENTREZID = character(), SYMBOL = character()))
  }
  suppressMessages(AnnotationDbi::select(
    org.Hs.eg.db, keys = accessions, keytype = "UNIPROT",
    columns = c("ENTREZID", "SYMBOL"))) %>%
    filter(!is.na(ENTREZID), !is.na(SYMBOL)) %>%
    distinct(UNIPROT, ENTREZID, SYMBOL) %>%
    add_count(UNIPROT) %>%
    filter(n == 1) %>%            # unambiguous accessions only
    select(-n) %>%
    rename(Substrate_Uniprot_ID = UNIPROT)
}

# 1) Collect all kinase/substrate names we’ll need to harmonize
# NB: both atlas files carry a `median_percentile` summary column that matches the
# `_(rank|percentile)$` pattern. It is not a kinase; leaving it in previously
# created a pseudo-kinase named "median" whose rows survived into the figure
# inputs as `Kinase = NA`. Exclude it explicitly here and in the pivot below.
kinases_cantley <- colnames(cantley_atlas %>% select(matches("_(rank|percentile)$"))) %>%
  str_replace("(_percentile|_rank)$","") %>% setdiff("median") %>% unique()
kinases_tyr     <- colnames(cantley_tyrosine %>% select(matches("_(rank|percentile)$"))) %>%
  str_replace("(_percentile|_rank)$","") %>% setdiff("median") %>% unique()
#kinases_psp     <- unique(phosphosite_plus$KINASE)
genes_hpa       <- unique(subcellular_localization$`Gene name`)
#kinases_pselm   <- unique(phospho_elm %>% filter(!str_detect(kinases, "group")) %>% pull(kinases))
#kinases_pn <- unique(phospho_networks$kinase)
kinases_kinet = kinet %>% filter(!is.na(`Kinase Name`)) %>% pull(`Kinase Name`) %>% unique()

subs_cantley    <- unique(c(cantley_atlas$Gene, cantley_tyrosine$Gene))
#subs_psp        <- unique(phosphosite_plus$SUBSTRATE)
#subs_pselm = bitr(unique(phospho_elm$acc), fromType="UNIPROT", toType="SYMBOL", OrgDb=org.Hs.eg.db) %>%
#  pull(SYMBOL)
#subs_pn = unique(phospho_networks$gene)
subs_kinet = kinet %>% filter(!is.na(`Substrate Name`)) %>% pull(`Substrate Name`) %>% unique()

# 2) Build symbol maps
symmap_kinase <- build_symbol_map(unique(c(kinases_cantley, kinases_tyr, genes_hpa, kinases_kinet)))
symmap_subs   <- build_symbol_map(unique(c(subs_cantley, genes_hpa, subs_kinet)))

# 3) Authoritative resolver for the specificity-atlas kinase column names.
#
# The atlases label kinase columns with Manning/CORAL-style short names
# ("P38A", "CK1D", "ABL", "P90RSK", and "*_TYR" variants for kinases profiled in
# the tyrosine atlas) rather than HGNC symbols. Resolving those through
# org.Hs.eg.db is unsafe in both directions: 51 of the 396 atlas names have no
# SYMBOL match at all (and were previously dropped silently), while several
# short names are ambiguous aliases -- "ABL" resolves to ABL1, MTTP and DSP-AS1,
# so a tyrosine-kinase profile was being attributed to a lipid-transfer protein
# and an antisense RNA. We resolve atlas names against the curated kinome table
# (CORAL.ID, then unambiguous Alias, then HGNC symbol), fall back to an explicit
# override list, and stop the build if anything is still unmapped.
atlas_kinase_overrides <- tibble::tribble(
  ~key,      ~Kinase,
  "MKK4",    "MAP2K4",
  "MKK6",    "MAP2K6",
  "MKK7",    "MAP2K7",
  "P90RSK",  "RPS6KA1",
  "ABL",     "ABL1",
  "ARG",     "ABL2",
  "ETK",     "BMX",
  # Manning/CORAL alpha-kinase numbering does not follow HGNC numbering:
  # CORAL "AlphaK3" is ALPK1 (an active kinase), not ALPK3 (a pseudokinase).
  "ALPHAK3", "ALPK1"
)

kinase_ref <- read_excel(here("data/Comprehensive_Kinase_Simple.xlsx"),
                         .name_repair = "unique_quiet")

# The curated kinome table is authoritative for the kinases it lists. org.Hs.eg.db
# is not, for a handful of them: it lists two Entrez IDs for the symbol "TEC", so
# build_symbol_map() discards it, and it no longer knows ICK / PRPF4B / STK19 /
# COL4A3BP under their old names. Left alone, those kinases reach the curated layer
# with no Entrez ID (or a different symbol from the atlas layer), which breaks every
# downstream join. Override the map for exactly the names the table lists.
kinome_symbol_map <- kinase_ref %>%
  transmute(input_symbol = as.character(Gene), ENTREZID = as.character(`Human Gene ID`)) %>%
  filter(!is.na(input_symbol), !is.na(ENTREZID)) %>%
  distinct(input_symbol, .keep_all = TRUE)
kinome_current_symbol <- AnnotationDbi::mapIds(
  org.Hs.eg.db, keys = kinome_symbol_map$ENTREZID, column = "SYMBOL",
  keytype = "ENTREZID", multiVals = "first")
kinome_symbol_map <- kinome_symbol_map %>%
  mutate(SYMBOL = unname(kinome_current_symbol[ENTREZID])) %>%
  filter(!is.na(SYMBOL))
kinome_map_changes <- symmap_kinase %>%
  inner_join(kinome_symbol_map, by = "input_symbol", suffix = c("", "_ref")) %>%
  filter(ENTREZID != ENTREZID_ref)
if (nrow(kinome_map_changes)) {
  message("Kinome table overrides org.Hs.eg.db for ", nrow(kinome_map_changes), " name(s): ",
          paste0(kinome_map_changes$input_symbol, " (", kinome_map_changes$ENTREZID, " -> ",
                 kinome_map_changes$ENTREZID_ref, ")", collapse = ", "))
}
symmap_kinase <- bind_rows(
  anti_join(symmap_kinase, kinome_symbol_map, by = "input_symbol"),
  kinome_symbol_map) %>%
  distinct()

kinase_key_map <- bind_rows(
  atlas_kinase_overrides %>% transmute(key = toupper(key), Kinase, pri = 0L),
  kinase_ref %>% transmute(key = toupper(str_trim(Gene)), Kinase = Gene, pri = 1L),
  kinase_ref %>% transmute(key = toupper(str_trim(as.character(CORAL.ID))),
                           Kinase = Gene, pri = 2L),
  kinase_ref %>%
    transmute(Alias = as.character(Alias), Kinase = Gene) %>%
    separate_rows(Alias, sep = "[,;|]\\s*") %>%
    transmute(key = toupper(str_trim(Alias)), Kinase, pri = 3L) %>%
    add_count(key) %>% filter(n == 1) %>% select(-n),
  # last resort: org.Hs.eg.db, but only via build_symbol_map(), which has
  # already discarded any alias resolving to more than one gene. This covers
  # widely-used trivial names (MEK1 -> MAP2K1, HER2 -> ERBB2, PERK -> EIF2AK3,
  # VEGFR2 -> KDR, ...) that the curated kinome table does not list.
  symmap_kinase %>% transmute(key = toupper(input_symbol), Kinase = SYMBOL, pri = 4L)
) %>%
  filter(!is.na(key), key != "", key != "NA", !is.na(Kinase)) %>%
  arrange(pri) %>%
  distinct(key, .keep_all = TRUE) %>%
  select(key, Kinase)

kinase_ref_ids <- kinase_ref %>%
  transmute(Kinase = Gene, ref_entrez = as.character(`Human Gene ID`)) %>%
  filter(!is.na(Kinase), !is.na(ref_entrez)) %>%
  distinct(Kinase, .keep_all = TRUE)

resolve_atlas_kinases <- function(atlas_names) {
  out <- tibble::tibble(Kinase_input = unique(na.omit(atlas_names))) %>%
    mutate(key = toupper(str_remove(Kinase_input, "_TYR$"))) %>%
    left_join(kinase_key_map, by = "key")
  unmapped <- out %>% filter(is.na(Kinase)) %>% pull(Kinase_input)
  if (length(unmapped)) {
    stop("Unmapped specificity-atlas kinase names (add to atlas_kinase_overrides): ",
         paste(sort(unmapped), collapse = ", "))
  }
  # Entrez ID: the curated kinome table's `Human Gene ID` is authoritative (it is
  # complete, and covers names such as TEC -- ambiguous in org.Hs.eg.db, which
  # lists two Entrez IDs for it -- and ICK/PRPF4B, which org.Hs.eg.db now knows
  # only under their current symbols CILK1/PRP4K). org.Hs.eg.db is the fallback
  # for names the table does not list.
  out <- out %>%
    left_join(kinase_ref_ids, by = "Kinase") %>%
    left_join(symmap_kinase %>% distinct(SYMBOL, ENTREZID),
              by = c("Kinase" = "SYMBOL")) %>%
    mutate(Kinase_entrez = coalesce(ref_entrez, ENTREZID)) %>%
    distinct(Kinase_input, .keep_all = TRUE)

  n_disagree <- sum(!is.na(out$ref_entrez) & !is.na(out$ENTREZID) &
                      out$ref_entrez != out$ENTREZID)
  if (n_disagree > 0) {
    message("WARNING: ", n_disagree, " atlas kinase(s) where the kinome table and ",
            "org.Hs.eg.db disagree on the Entrez ID; the kinome table was used.")
  }
  if (anyNA(out$Kinase_entrez)) {
    stop("Atlas kinase(s) with no Entrez ID: ",
         paste(sort(out$Kinase_input[is.na(out$Kinase_entrez)]), collapse = ", "))
  }

  # Report every gene under its current HGNC symbol, so that the atlas layer and
  # the curated layer (which is resolved through org.Hs.eg.db) always agree on the
  # symbol for a given gene -- otherwise the same kinase-substrate pair can appear
  # twice under an old and a new symbol (e.g. ICK / CILK1, PRPF4B / PRP4K).
  current_symbol <- AnnotationDbi::mapIds(
    org.Hs.eg.db, keys = unique(out$Kinase_entrez), column = "SYMBOL",
    keytype = "ENTREZID", multiVals = "first")
  out %>%
    mutate(Kinase = dplyr::coalesce(unname(current_symbol[Kinase_entrez]), Kinase)) %>%
    transmute(Kinase_input, Kinase, Kinase_entrez)
}

atlas_kinase_map <- resolve_atlas_kinases(c(kinases_cantley, kinases_tyr))
message("Resolved ", nrow(atlas_kinase_map), " specificity-atlas kinase names to ",
        dplyr::n_distinct(atlas_kinase_map$Kinase), " distinct genes.")

# # 3) Normalize UniProt IDs early
# phosphosite_plus <- phosphosite_plus %>%
#   mutate(SUB_ACC_ID = norm_uniprot(SUB_ACC_ID))

```

### Weights and Helper Functions
```{r}
# ── Tunable thresholds ─────────────────────────────────────────────────────────
# Motif support: a site must rank in the top 10% of all phosphosites scored for
# a given kinase, i.e. at or above the 90th percentile. This is the value stated
# in the manuscript and matches the usage in Johnson et al. 2023 and
# Yaron-Barir et al. 2024. Sensitivity across the 80th-99th percentiles is
# reported separately (docs/revision_analyses.qmd).
motif_percentile_cutoff <- 90

# Phosphosite functional score: Ochoa et al. 2020 report that at a 0.5 cut-off
# "around 10% of the phosphoproteome remains with 50% of the true positive
# phosphosites recovered, suggesting that this is a useful cut-off for an
# initial prioritization". Sensitivity across 0.3-0.7 is reported separately.
functional_score_cutoff <- 0.5

# Relaxed motif tier, released as a separate supplementary layer. Kinases whose
# specificity is dominated by docking interactions rather than by the catalytic
# cleft (SH2/PTB-mediated recruitment, MAPK D- and DEF-motifs) can phosphorylate
# sites that score poorly against their linear motif, so a single stringent tier
# structurally excludes them. This layer is clearly labelled and is NOT part of
# the high-confidence set.
relaxed_percentile_cutoff <- 75

# ── Tunable weights ────────────────────────────────────────────────────────────
# PSP evidence weights (you can re-tune these later if desired)
w_invitro <- 0.5
w_invivo  <- 1.0

# ── Helpers ────────────────────────────────────────────────────────────────────
safe_scale <- function(x) {
  if (all(is.na(x))) return(rep(NA_real_, length(x)))
  s <- sd(x, na.rm = TRUE)
  m <- mean(x, na.rm = TRUE)
  if (is.na(s) || s == 0) return((x - m) * 0)
  (x - m) / s
}

make_site_key <- function(entrez, site_chr) {
  site_chr <- str_trim(site_chr)
  site    <- str_extract(site_chr, "^[STY]\\d+$")
  res     <- str_sub(site, 1, 1)
  pos     <- suppressWarnings(as.integer(str_sub(site, 2)))
  paste(entrez, res, pos, sep=":")
}

```

# Process Cantley (Ser/Thr + Tyr)

```{r}

# Motif priors from Cantley

# Build site-bearing Cantley features (no combined score)
cantley_serthr_long <- cantley_atlas %>%
  pivot_longer(cols = matches("_(rank|percentile)$") & !matches("^median_"),
               names_to = "Kinase", values_to = "Value") %>%
  mutate(Metric = if_else(str_detect(Kinase, "_rank$"), "Rank", "Percentile"),
         Kinase = str_replace(Kinase, "(_percentile|_rank)$",""))

cantley_tyr_long <- cantley_tyrosine %>%
  pivot_longer(cols = matches("_(rank|percentile)$") & !matches("^median_"),
               names_to = "Kinase", values_to = "Value") %>%
  mutate(Metric = if_else(str_detect(Kinase, "_rank$"), "Rank", "Percentile"),
         Kinase = str_replace(Kinase, "(_percentile|_rank)$",""))

# Substrate identity: prefer the atlas gene label when it resolves to exactly
# one gene, otherwise fall back to the row's UniProt accession. This is what
# stops obsolete/ambiguous aliases (e.g. "SMAP") from being fanned out across
# several unrelated substrates.
subs_by_uniprot <- resolve_by_uniprot(unique(c(
  cantley_atlas$`Uniprot Primary Accession`,
  cantley_tyrosine$`Uniprot Primary Accession`
)))

cantley_feats <- bind_rows(cantley_serthr_long, cantley_tyr_long) %>%
  pivot_wider(names_from = Metric, values_from = Value) %>%
  rename(Kinase_input = Kinase, Substrate_input = Gene) %>%
  mutate(Substrate_Uniprot_ID = norm_uniprot(`Uniprot Primary Accession`)) %>%
  # kinase identity from the authoritative atlas resolver (never a raw ALIAS hit)
  left_join(atlas_kinase_map, by = "Kinase_input") %>%
  # substrate identity: symbol first, UniProt accession as fallback
  left_join(symmap_subs, by = c("Substrate_input" = "input_symbol")) %>%
  left_join(subs_by_uniprot, by = "Substrate_Uniprot_ID",
            suffix = c("", ".up")) %>%
  mutate(
    Substrate        = coalesce(SYMBOL, SYMBOL.up),
    Substrate_entrez = coalesce(ENTREZID, ENTREZID.up)
  ) %>%
  transmute(
    Substrate_Uniprot_ID,
    Substrate, Substrate_entrez, Substrate_input,
    Kinase, Kinase_entrez, Kinase_input,
    Phosphosite = Phosphosite,
    Percentile, Rank, promiscuity_index
  ) %>%
  mutate(
    Phosphosite = str_trim(Phosphosite),
    Site        = str_extract(Phosphosite, "^[STY]\\d+$"),
    residue     = str_sub(Site, 1, 1),
    position    = suppressWarnings(as.integer(str_sub(Site, 2))),
    zP  = as.numeric(scale(Percentile)),
    zR  = as.numeric(scale(-Rank)),
    zPI = as.numeric(scale(-log1p(promiscuity_index)))
  ) %>%
  distinct()

# quick sanity
cantley_feats %>%
  summarise(
    n_pairs      = n(),
    n_kinases    = n_distinct(Kinase),
    n_substrates = n_distinct(Substrate_Uniprot_ID)
  )

promiscuity_index_cutoff = quantile(cantley_feats$promiscuity_index, 0.5, na.rm=TRUE)

# cantley_proximity = cantley_feats %>%
#   filter(!is.na(Percentile)) %>%
#   filter(promiscuity_index < promiscuity_index_cutoff)

# #quick look at percentile distribution after promiscuity filtering
# ggplot(cantley_proximity, aes(x=Percentile)) +
#   geom_histogram() +
#   #add quantile lines and labels
#   geom_vline(xintercept = quantile(cantley_proximity$Percentile, 0.75, na.rm=TRUE), linetype="dashed", color="red") +
#   geom_vline(xintercept = quantile(cantley_proximity$Percentile, 0.9, na.rm=TRUE), linetype="dashed", color="blue") +
#   geom_vline(xintercept = quantile(cantley_proximity$Percentile, 0.99, na.rm=TRUE), linetype="dashed", color="green") +
#   annotate("text", x=quantile(cantley_proximity$Percentile, 0.75, na.rm=TRUE), y=50, label="75%", color="red", angle=90, vjust=-0.5) +
#   annotate("text", x=quantile(cantley_proximity$Percentile, 0.9, na.rm=TRUE), y=50, label="90%", color="blue", angle=90, vjust=-0.5) +
#   annotate("text", x=quantile(cantley_proximity$Percentile, 0.99, na.rm=TRUE), y=50, label="99%", color="green", angle=90, vjust=-0.5) +
#   theme_pubr() +
#   labs(title="Cantley Atlas Percentile Distribution after Promiscuity Filtering",
#        x="Percentile", y="Count")

cantley_top <- cantley_feats %>%
  filter(!is.na(Percentile), Percentile >= motif_percentile_cutoff) %>%
  filter(promiscuity_index < promiscuity_index_cutoff) %>%
  # standardize a site key for filtering
  mutate(site_key = make_site_key(Substrate_entrez, Site)) %>%
  distinct()


```

## Phosphositeplus, PhosphoELM, Phosphonetwork processing

```{r eval=FALSE}
## Phsophositeplus processing ----------------------------------
phosphosite_processed <- phosphosite_plus %>%
  rename(
    Kinase               = KINASE,
    Substrate       = SUBSTRATE,
    Substrate_Uniprot_ID = SUB_ACC_ID,
    Kinase_Uniprot_ID = KIN_ACC_ID) %>%
  mutate(
    score_raw = (if_else(!is.na(IN_VIVO_RXN),  w_invivo,  0)) +
                (if_else(!is.na(IN_VITRO_RXN), w_invitro, 0))
  ) %>%
  distinct() %>%
  filter(KIN_ORGANISM == "human", SUB_ORGANISM == "human") %>%
  left_join(symmap_kinase, by = c("Kinase" = "input_symbol")) %>%
  rename(Kinase_input = Kinase, Kinase = SYMBOL, Kinase_entrez = ENTREZID) %>%
  left_join(symmap_subs,   by = c("Substrate" = "input_symbol")) %>%
  rename(Substrate_input = Substrate, Substrate = SYMBOL, Substrate_entrez = ENTREZID) %>%
  mutate(Substrate = coalesce(Substrate, Substrate_input))

psp_pairs <- phosphosite_processed %>%
  select(Kinase, Kinase_entrez, Kinase_Uniprot_ID, Substrate, Substrate_entrez, Substrate_Uniprot_ID) %>%
  distinct()

psp_sites <- phosphosite_processed %>%
  rename(Site_raw = SUB_MOD_RSD) %>%
  separate_rows(Site_raw, sep = "[,;|]") %>%
  mutate(Site_raw = str_trim(Site_raw),
         Site     = str_extract(Site_raw, "^[STY]\\d+$")) %>%
  filter(!is.na(Site)) %>%
  mutate(residue  = str_sub(Site, 1, 1),
         position = as.integer(str_sub(Site, 2))) %>%
  distinct(Substrate, Substrate_entrez, Substrate_Uniprot_ID, Site, residue, position)


## PhosphoELM processing ----------------------------------

phosphoelm_sub_symbols = bitr(unique(phospho_elm$acc), fromType="UNIPROT", toType="SYMBOL", OrgDb=org.Hs.eg.db)
phosphoelm_kinase_uniprot = bitr(unique(phospho_elm$kinases), fromType="SYMBOL", toType="UNIPROT", OrgDb=org.Hs.eg.db) %>%
  #select first uniprot symbol per kinase
  group_by(SYMBOL) %>%
  slice(1) %>%
  ungroup()

phospho_elm_processed = phospho_elm %>%
  rename(Substrate_Uniprot_ID = acc,
        Kinase = kinases) %>%
  filter(!is.na(Kinase)) %>%
  filter(!str_detect(Kinase, "group")) %>%
  filter(species == "Homo sapiens") %>%
  select(-species) %>%
  left_join(phosphoelm_sub_symbols, by = c("Substrate_Uniprot_ID" = "UNIPROT")) %>%
  rename(Substrate = SYMBOL) %>%
  mutate(Substrate = coalesce(Substrate, Substrate_Uniprot_ID)) %>%
  left_join(symmap_subs, by = c("Substrate" = "input_symbol")) %>%
  rename(Substrate_input = Substrate, Substrate = SYMBOL, Substrate_entrez = ENTREZID) %>%
  mutate(Substrate = coalesce(Substrate, Substrate_input)) %>%
  left_join(symmap_kinase, by = c("Kinase" = "input_symbol")) %>%
  rename(Kinase_input = Kinase, Kinase = SYMBOL, Kinase_entrez = ENTREZID) %>%
  mutate(Kinase = coalesce(Kinase, Kinase_input)) %>%
  left_join(phosphoelm_kinase_uniprot, by = c("Kinase" = "SYMBOL"))

phosphoelm_pairs = phospho_elm_processed %>%
  distinct(Kinase, Kinase_entrez, Substrate, Substrate_entrez, Substrate_Uniprot_ID)

phospho_elm_sites <- phospho_elm_processed %>%
  rename(residue = code,
         position = position) %>%
  distinct(Substrate, Substrate_entrez, Substrate_Uniprot_ID, Site = paste0(residue, position), residue, position)

```

```{r eval=FALSE}
# ## Phosphonetwork Processing
# fasta_path = here("data/UP000005640_9606.fasta")
#   # Identify header lines (start with ">")
#   header_idx <- grep("^>", phospho_networks)
  
#   # Add an end marker at the last line
#   header_idx <- c(header_idx, length(phospho_networks) + 1)

#   # Iterate over each block
#   entries <- map_dfr(seq_along(header_idx[-length(header_idx)]), function(i) {
#     start <- header_idx[i]
#     end <- header_idx[i+1] - 1

#     header <- sub("^>", "", phospho_networks[start])
#     fields <- strsplit(header, "\t")[[1]]
#     kinase <- fields[1]
#     residue <- fields[2]
#     count <- as.integer(fields[3])

#     peptides <- phospho_networks[(start+1):end]
#     tibble(kinase = kinase,
#            residue = residue,
#            count = count,
#            peptide = peptides)
#   })

# # -------- 1) Reader for the FASTA-like PhosphoNetworks file --------
# read_phosphonetworks <- function(path) {
#   lines <- readLines(path)
#   hdr_idx <- grep("^>", lines)
#   hdr_idx <- c(hdr_idx, length(lines) + 1)

#   map_dfr(seq_along(hdr_idx[-length(hdr_idx)]), function(i) {
#     start <- hdr_idx[i]; end <- hdr_idx[i+1] - 1
#     header <- sub("^>", "", lines[start])
#     fields <- strsplit(header, "\t")[[1]]
#     stopifnot(length(fields) >= 3)
#     tibble(
#       kinase  = fields[1],
#       residue_class = fields[2],           # "ST" or "Y"
#       n_reported = suppressWarnings(as.integer(fields[3])),
#       peptide_raw = lines[(start+1):end]
#     )
#   }) %>%
#     unnest(peptide_raw) %>%
#     filter(!is.na(peptide_raw), peptide_raw != "")
# }

# # -------- 2) Helpers for peptide cleaning & center residue logic --------
# # Keep AA letters only, uppercase; return cleaned string
# clean_peptide <- function(x) {
#   x <- toupper(x)
#   x <- gsub("[^ACDEFGHIKLMNPQRSTVWY]", "", x)  # drop hyphens/unknowns/spaces
#   x
# }

# # Find the "central" residue defined by the center of the *raw* string,
# # then map that index into the cleaned string (ignoring hyphens etc.)
# central_index_info <- function(raw_seq) {
#   c_raw <- ceiling(nchar(raw_seq)/2)
#   chars <- strsplit(toupper(raw_seq), "")[[1]]
#   aa_set <- strsplit("ACDEFGHIKLMNPQRSTVWY","")[[1]]
#   aa_idx <- which(chars %in% aa_set)

#   if (length(aa_idx) == 0L) return(NA_integer_)   # <- important guard

#   if (!(c_raw %in% aa_idx)) {
#     nearest <- aa_idx[which.min(abs(aa_idx - c_raw))]
#     c_raw <- nearest
#   }
#   as.integer(match(c_raw, aa_idx))
# }

# # Check that the center AA matches expected class
# center_matches_class <- function(clean_seq, c_idx, class_str) {
#   if (is.na(c_idx) || c_idx < 1 || c_idx > nchar(clean_seq)) return(FALSE)
#   aa <- substr(clean_seq, c_idx, c_idx)
#   if (class_str == "Y") {
#     return(aa == "Y")
#   } else { # "ST"
#     return(aa %in% c("S","T"))
#   }
# }

# # -------- 3) Parse UniProt FASTA headers (sp|...|ENTRY OS=... GN=... etc.) --------
# parse_uniprot_header <- function(h) {
#   # typical header: >sp|Q9Y5Q8|TF3C5_HUMAN ... OS=Homo sapiens GN=GTF3C5 PE=1 SV=2
#   reviewed <- ifelse(grepl("^sp\\|", h), TRUE, FALSE)
#   acc <- sub("^..\\|([^|]+)\\|.*$", "\\1", h)
#   entry <- sub("^..\\|[^|]+\\|([^ ]+).*$", "\\1", h)
#   gene <- ifelse(grepl("\\bGN=[^ ]+", h),
#                  sub(".*\\bGN=([^ ]+).*", "\\1", h),
#                  NA_character_)
#   tibble(accession = acc, entry = entry, gene = gene, reviewed = reviewed)
# }

# map_one <- function(pep, c_idx, proteome) {
#   # Guard against NA/empty
#   if (is.na(c_idx) || nchar(pep) == 0L) return(tibble())

#   hits <- vmatchPattern(AAString(pep), proteome, fixed = TRUE)
#   if (length(hits) == 0L) return(tibble())

#   subj_idx <- which(lengths(hits) > 0L)
#   # Build rows for all subjects with any hit
#   rows <- lapply(subj_idx, function(si) {
#     st <- start(hits[[si]])
#     en <- end(hits[[si]])
#     # coerce to plain integer vectors
#     st <- as.integer(st)
#     en <- as.integer(en)
#     # center_pos per match in this subject
#     cp <- st + (as.integer(c_idx) - 1L)
#     tibble(seq_index = si, start = st, end = en, center_pos = cp)
#   })
#   bind_rows(rows)
# }

# # -------- 4) Map peptides to the proteome --------
# map_peptides_to_proteome <- function(df_pep, proteome_fasta, species_regex = "Homo sapiens",
#                                      prefer_reviewed = TRUE,
#                                      require_unique_in_proteome = TRUE,
#                                      require_unique_in_protein = TRUE) {
#   proteome <- readAAStringSet(proteome_fasta)
#   headers  <- names(proteome)

#   # keep only human (defensive)
#   keep_idx <- grepl(species_regex, headers)
#   proteome <- proteome[keep_idx]
#   headers  <- headers[keep_idx]

#   # parse UniProt meta
#   meta <- purrr::map_dfr(headers, parse_uniprot_header) |>
#     mutate(seq_index = row_number())

#   # Pre-clean peptides & center
#   prep <- df_pep |>
#     mutate(
#       pep_clean         = map_chr(peptide_raw, clean_peptide),
#       center_clean_idx  = map_int(peptide_raw, central_index_info),
#       center_ok         = mapply(center_matches_class, pep_clean, center_clean_idx, residue_class),
#       pep_len           = nchar(pep_clean)
#     ) |>
#     filter(pep_len > 0, center_ok) |>
#     distinct(kinase, residue_class, peptide_raw, pep_clean, center_clean_idx, .keep_all = TRUE) |>
#     mutate(pid = row_number())   # stable peptide id

#   # Map each peptide independently, keep pid with results
#   mapped <- prep |>
#     select(pid, kinase, residue_class, peptide_raw, pep_clean, center_clean_idx) |>
#     pmap_dfr(function(pid, kinase, residue_class, peptide_raw, pep_clean, center_clean_idx) {
#       out <- map_one(pep_clean, center_clean_idx, proteome)
#       if (nrow(out) == 0L) return(tibble())
#       out |>
#         mutate(pid = pid, kinase = kinase, residue_class = residue_class,
#                peptide_raw = peptide_raw, pep_clean = pep_clean)
#     })

#   if (nrow(mapped) == 0L) {
#     message("No peptide mapped to the proteome.")
#     return(tibble())
#   }

#   # Join UniProt meta, verify residue at center, label site
#   mapped2 <- mapped |>
#     left_join(meta, by = "seq_index") |>
#     mutate(center_letter = mapply(function(si, pos) {
#       if (is.na(pos) || pos < 1L) return(NA_character_)
#       as.character(subseq(proteome[[si]], start = pos, end = pos))
#     }, seq_index, center_pos)) |>
#     filter(!is.na(center_letter)) |>
#     mutate(center_ok_proteome = case_when(
#       residue_class == "Y"  ~ center_letter == "Y",
#       residue_class == "ST" ~ center_letter %in% c("S","T"),
#       TRUE ~ FALSE
#     )) |>
#     filter(center_ok_proteome) |>
#     mutate(site = paste0(center_letter, center_pos),
#            source = "PhosphoNetworks_motif→UniProt_map")

#   # Ambiguity controls
#   if (prefer_reviewed) {
#     mapped2 <- mapped2 |>
#       group_by(pid) |>
#       mutate(any_reviewed = any(reviewed)) |>
#       ungroup() |>
#       filter(!any_reviewed | reviewed) |>
#       select(-any_reviewed)
#   }

#   if (require_unique_in_proteome) {
#     mapped2 <- mapped2 |>
#       group_by(pid) |>
#       mutate(n_hits_proteome = n()) |>
#       ungroup() |>
#       filter(n_hits_proteome == 1L)
#   }

#   if (require_unique_in_protein) {
#     mapped2 <- mapped2 |>
#       group_by(pid, accession) |>
#       mutate(n_hits_this_protein = n()) |>
#       ungroup() |>
#       filter(n_hits_this_protein == 1L)
#   }

#   mapped2 |>
#     select(kinase, residue_class, peptide = pep_clean, peptide_raw,
#            accession, entry, gene, reviewed,
#            site, residue_letter = center_letter, position = center_pos,
#            source) |>
#     arrange(kinase, gene, position)
# }

# # -------- 5) DRIVER --------
# # paths (change to your actual files)
# pn_path   <- here("data/motifSite.csv")        # your motif file
# fasta_path <- here("data/UP000005640_9606.fasta")   # Swiss-Prot HUMAN proteome FASTA

# pn_df <- read_phosphonetworks(pn_path)

# # ks_sites <- map_peptides_to_proteome(
# #   df_pep = pn_df,
# #   proteome_fasta = fasta_path,
# #   species_regex = "Homo sapiens",
# #   prefer_reviewed = TRUE,
# #   require_unique_in_proteome = TRUE,   # strong ambiguity guard
# #   require_unique_in_protein = TRUE
# # )

# write_csv(ks_sites, here("data/phospho_networks_kinase_substrate_sites.csv"))

```

```{r eval=FALSE}
pn_kinase_uniprot = bitr(phospho_networks$kinase, fromType = "SYMBOL", toType = "UNIPROT", OrgDb = org.Hs.eg.db)

phospho_networks_processed = phospho_networks %>%
  rename(Kinase = kinase, Substrate_Uniprot_ID = accession, Substrate = gene) %>%
  left_join(symmap_kinase, by = c("Kinase" = "input_symbol")) %>%
  rename(Kinase_entrez = ENTREZID, Kinase_input = Kinase, Kinase = SYMBOL) %>%
  mutate(Kinase = coalesce(Kinase, Kinase_input)) %>%
  distinct() %>%
  left_join(symmap_subs, by = c("Substrate" = "input_symbol")) %>%
  rename(Substrate_entrez = ENTREZID, Substrate_input = Substrate, Substrate = SYMBOL) %>%
  mutate(Substrate = coalesce(Substrate, Substrate_input)) %>%
  left_join(pn_kinase_uniprot, by = c("Kinase" = "SYMBOL")) %>%
  rename(Kinase_Uniprot_ID = UNIPROT)

phospho_networks_pairs = phospho_networks_processed %>%
  distinct(Kinase, Kinase_Uniprot_ID, Kinase_entrez, Substrate, Substrate_entrez, Substrate_Uniprot_ID)

phospho_networks_sites = phospho_networks_processed %>%
  select(Substrate, Substrate_Uniprot_ID, Site = site, residue = residue_letter, position) %>%
  distinct()

```

## Kinet processing
```{r}

kinet_processed = kinet %>%
  rename(
    Kinase_Uniprot_ID = Kinase,
    Kinase = `Kinase Name`,
    Substrate_Uniprot_ID = Substrate,
    Substrate = `Substrate Name`,
    Source = `Source Database`,
    Reference_PMID = `Reference (PMID)`
  ) %>%
  filter(!is.na(Kinase), !is.na(Substrate)) %>%
  left_join(symmap_kinase, by = c("Kinase" = "input_symbol")) %>%
  rename(Kinase_entrez = ENTREZID, Kinase_input = Kinase, Kinase = SYMBOL) %>%
  mutate(Kinase = coalesce(Kinase, Kinase_input)) %>%
  left_join(symmap_subs, by = c("Substrate" = "input_symbol")) %>%
  rename(Substrate_entrez = ENTREZID, Substrate_input = Substrate, Substrate = SYMBOL) %>%
  mutate(Substrate = coalesce(Substrate, Substrate_input)) %>%
  select(-Kinase_input, -Substrate_input) %>%
  select(Kinase, Kinase_entrez, Kinase_Uniprot_ID,
         Substrate, Substrate_entrez, Substrate_Uniprot_ID, Site,
         Source, Evidence, Reference_PMID) %>%
  distinct()

```

## Ochoa functional score processing

```{r}

ochoa_fs_processed = ochoa_fs %>%
  rename(Substrate_Uniprot_ID = uniprot)

# #quick look at functional score distribution
# ochoa_fs_processed %>%
#   ggplot(aes(x=functional_score)) +
#   geom_histogram(bins=50) +
#   geom_vline(xintercept=0.5, color="red", linetype="dashed")

```

## Aggregated Kinase Substrate dataset
```{r eval=FALSE}
ks_all_raw <- bind_rows(psp_pairs %>% mutate(Source = "PhosphoSitePlus"),
  phosphoelm_pairs %>% mutate(Source = "PhosphoELM"),
  phospho_networks_pairs %>% mutate(Source = "PhosphoNetworks")) %>%
  distinct() %>%
  filter(!is.na(Kinase) & !is.na(Substrate))  # drop unknowns

# 2) Unique gene-level pairs (no UniProt columns yet)
pairs_gene <- ks_all_raw %>%
  select(Kinase, Kinase_entrez, Substrate, Substrate_entrez, Source) %>%
  distinct()

# 3) Choose ONE UniProt per Kinase (most frequent non-NA; tie -> smallest ID)
kinase_choice <- ks_all_raw %>%
  filter(!is.na(Kinase_Uniprot_ID), Kinase_Uniprot_ID != "") %>%
  count(Kinase, Kinase_entrez, Kinase_Uniprot_ID, name = "n") %>%
  arrange(Kinase, Kinase_entrez, desc(n), Kinase_Uniprot_ID) %>%
  group_by(Kinase, Kinase_entrez) %>%
  slice(1) %>%
  ungroup() %>%
  select(Kinase, Kinase_entrez, Kinase_Uniprot_ID)

# 4) Choose ONE UniProt per Substrate (same rule)
substrate_choice <- ks_all_raw %>%
  filter(!is.na(Substrate_Uniprot_ID), Substrate_Uniprot_ID != "") %>%
  count(Substrate, Substrate_entrez, Substrate_Uniprot_ID, name = "n") %>%
  arrange(Substrate, Substrate_entrez, desc(n), Substrate_Uniprot_ID) %>%
  group_by(Substrate, Substrate_entrez) %>%
  slice(1) %>%
  ungroup() %>%
  select(Substrate, Substrate_entrez, Substrate_Uniprot_ID)

# 5) Reconstruct the final table with one UniProt per gene
ks_all <- pairs_gene %>%
  left_join(kinase_choice,   by = c("Kinase",   "Kinase_entrez")) %>%
  left_join(substrate_choice, by = c("Substrate","Substrate_entrez"))


ks_sites_all = psp_sites %>%
  mutate(Source = "PhosphoSitePlus") %>%
  bind_rows(phospho_elm_sites %>% mutate(Source = "PhosphoELM")) %>%
  bind_rows(phospho_networks_sites %>% mutate(Source = "PhosphoNetworks")) %>%
  distinct() %>%
  group_by(Source, Substrate, Substrate_entrez, Site, residue, position) %>%
  summarise(Substrate_Uniprot_ID = first(na.omit(Substrate_Uniprot_ID)), .groups="drop") %>%
  ungroup()

```



```{r}

# --- 0) Helper: make a site key like "P12345:S:241"

# =========================
# STEP 1: Global 90th percentile filter
# =========================

# #qucikly visualize promosicuity distribution
# cantley_feats %>%
#   filter(Percentile > 90) %>%
#   filter(!is.na(promiscuity_index)) %>%
#   ggplot(aes(x=promiscuity_index)) +
#   geom_histogram(bins=50) +
#   #add lines at quantiles and label
#   geom_vline(xintercept=quantile(cantley_feats$promiscuity_index, 0.25, na.rm=TRUE), color="green", linetype="dashed") +
#   geom_vline(xintercept=quantile(cantley_feats$promiscuity_index, 0.5, na.rm=TRUE), color="blue", linetype="dashed") +
#   geom_vline(xintercept=quantile(cantley_feats$promiscuity_index, 0.75, na.rm=TRUE), color="purple", linetype="dashed") +
#   geom_vline(xintercept=quantile(cantley_feats$promiscuity_index, 0.9, na.rm=TRUE), color="red", linetype="dashed") +
#   theme_pubr()

promiscuity_index_cutoff = quantile(cantley_feats$promiscuity_index, 0.5, na.rm=TRUE)

cantley_top <- cantley_feats %>%
  filter(!is.na(Percentile), Percentile >= motif_percentile_cutoff) %>%
  filter(promiscuity_index < promiscuity_index_cutoff) %>%
  # standardize a site key for filtering
  mutate(site_key = make_site_key(Substrate_entrez, Site)) %>%
  distinct()

# =========================
# STEP 2: HPA co-localization filter
# (re-using your HPA parsing approach; assumes subcellular_localization and symmaps exist)
# =========================

hpa_entrez = bitr(unique(subcellular_localization$`Gene name`), fromType="SYMBOL", toType="ENTREZID", OrgDb=org.Hs.eg.db)

hpa_long <- subcellular_localization %>%
  select(-Gene) %>%
  rename(Gene = `Gene name`) %>%
  pivot_longer(
    cols = c(contains("Approved"), contains("Supported"), contains("Enhanced")),
    names_to = "tier", values_to = "locs"
  ) %>%
  filter(!is.na(locs)) %>%
  separate_rows(locs, sep = ";") %>%
  mutate(locs = str_trim(locs)) %>%
  filter(locs != "") %>%
  left_join(
    hpa_entrez,
    by = c("Gene" = "SYMBOL")
  ) %>%
  transmute(Gene_entrez = ENTREZID, locs) %>%
  # HPA genes that do not map to an Entrez ID must be dropped: an NA key would
  # otherwise match any kinase or substrate that itself lacks an Entrez ID and
  # manufacture a spurious shared compartment.
  filter(!is.na(Gene_entrez)) %>%
  distinct()

# long forms limited to needed genes
genes_needed <- unique(c(cantley_top$Kinase_entrez, cantley_top$Substrate_entrez))
hpa_long_small <- hpa_long %>% filter(Gene_entrez %in% genes_needed)

kin_long <- hpa_long_small %>% transmute(loc = locs, Kinase_entrez = Gene_entrez) %>% distinct()
sub_long <- hpa_long_small %>% transmute(loc = locs, Substrate_entrez = Gene_entrez) %>% distinct()

# KS keys in cantley_top
ks_eval_keys <- unique(paste(cantley_top$Kinase_entrez, cantley_top$Substrate_entrez, sep="|"))
ks_eval_dt   <- data.table(ks_key = ks_eval_keys)

# cross by location -> candidate co-loc pairs
kin_dt <- as.data.table(kin_long); setkey(kin_dt, loc)
sub_dt <- as.data.table(sub_long); setkey(sub_dt, loc)

# The join is on `loc`, and kin_long/sub_long are distinct on (loc, gene), so the
# number of rows per kinase-substrate key is exactly the number of HPA
# compartments the two proteins share. Keep that count, not just the boolean:
# it is reported in the released tables and feeds the composite score, and it
# lets users distinguish a pair sharing one compartment from one sharing six.
coloc_pairs <- kin_dt[sub_dt, allow.cartesian = TRUE][
  , .(ks_key = paste(Kinase_entrez, Substrate_entrez, sep="|"))][
  , .(n_shared_compartments = .N), by = ks_key]

# reduce to only pairs present in cantley_top
setkey(coloc_pairs, ks_key); setkey(ks_eval_dt, ks_key)
coloc_counts <- coloc_pairs[ks_eval_dt, nomatch = 0L]

# flag co-loc in cantley_top
cantley_top_dt <- as.data.table(cantley_top)
cantley_top_dt[, ks_key := paste(Kinase_entrez, Substrate_entrez, sep="|")]
cantley_top_dt[coloc_counts, n_shared_compartments := i.n_shared_compartments, on = "ks_key"]
cantley_top_dt[is.na(n_shared_compartments), n_shared_compartments := 0L]
cantley_top_dt[, coloc_ok := n_shared_compartments > 0L]

# =========================
# STEP 3: Annotate curated + Ochoa, then filter with coloc_ok
# =========================

# (A) Curated-site membership (KiNet)
ks_sites_norm <- kinet_processed %>%
  mutate(
    Site = if ("Site" %in% names(.)) Site else paste0(residue, position),
    site_key = make_site_key(Substrate_entrez, Site)
  ) %>%
  mutate(residue = str_sub(Site, 1, 1),
         position = suppressWarnings(as.integer(str_sub(Site, 2)))) %>%
  transmute(Substrate_entrez, residue, position, site_key) %>%
  distinct()

curated_keys <- unique(ks_sites_norm$site_key)
cantley_top_dt[, curated_site := !is.na(fmatch(site_key, curated_keys))]

# #quick look at functional score distribution
# ochoa_fs_processed %>%
#   ggplot(aes(x=functional_score)) +
#   geom_histogram(bins=50) +
#   #add a line for quartiles and label
#   geom_vline(xintercept=quantile(ochoa_fs_processed$functional_score, 0.25, na.rm=TRUE), color="green", linetype="dashed") +
#   geom_vline(xintercept=quantile(ochoa_fs_processed$functional_score, 0.5, na.rm=TRUE), color="blue", linetype="dashed") +
#   geom_vline(xintercept=quantile(ochoa_fs_processed$functional_score, 0.75, na.rm=TRUE), color="purple", linetype="dashed") +
#   geom_vline(xintercept=0.5, color="red", linetype="dashed")

cutoff <- functional_score_cutoff

# (B) Ochoa functional score (>= functional_score_cutoff) -- join by UniProt + position
# Make sure both sides use normalized UniProt and integer position
ochoa_fs_norm <- ochoa_fs %>%
  transmute(
    Substrate_Uniprot_ID = norm_uniprot(uniprot),
    position = as.integer(position),
    functional_score
  ) %>%
  distinct()

cantley_top_annot <- as_tibble(cantley_top_dt) %>%
  mutate(position = as.integer(position)) %>%
  left_join(
    ochoa_fs_norm,
    by = c("Substrate_Uniprot_ID", "position")
  ) %>%
  mutate(ochoa_high = !is.na(functional_score) & functional_score >= cutoff)

# (B2) Evidence-layer contingency table.
# All three filter indicators are available on a single object here, before any
# of them has been applied, so this is the natural place to record how much
# information each layer carries and how far they overlap. Feeds the
# orthogonality / mutual-information analysis and the filter-rejection UpSet.
filter_contingency <- cantley_top_annot %>%
  count(coloc_ok, curated_site, ochoa_high, name = "n_site_kinase_pairs")
write_csv(filter_contingency, here("results/filter_evidence_contingency.csv"))
print(filter_contingency)

# (B3) Continuous confidence score.
# The binary flags are retained for backwards compatibility, but each evidence
# layer is also mapped through a monotonic, continuously differentiable
# transform so that a site scoring just below a threshold is down-weighted
# rather than discarded outright. Each component is reported separately so users
# can reweight them; the default is an unweighted mean.
soft_step <- function(x, midpoint, steepness) 1 / (1 + exp(-steepness * (x - midpoint)))

motif_steepness      <- 0.15   # transition spans roughly +/- 20 percentile points
functional_steepness <- 12     # transition spans roughly +/- 0.25 score units
compartment_scale    <- 2      # compartments at which localization evidence saturates

cantley_top_annot <- cantley_top_annot %>%
  mutate(
    motif_evidence        = soft_step(Percentile, motif_percentile_cutoff, motif_steepness),
    functional_evidence   = soft_step(coalesce(functional_score, 0),
                                      functional_score_cutoff, functional_steepness),
    localization_evidence = 1 - exp(-n_shared_compartments / compartment_scale),
    kinasecanvas_score    = (motif_evidence + functional_evidence + localization_evidence) / 3
  )

# (C) Final filter: co-localized AND (curated OR Ochoa-high).
# Note this is deliberately a disjunction on the two site-level criteria: a site
# already curated in PhosphoSitePlus/EPSD/iPTMnet is retained regardless of its
# functional score, because curation is itself direct evidence of relevance.
cantley_filtered_dt <- as.data.table(cantley_top_annot)[
  coloc_ok == TRUE & (curated_site == TRUE | ochoa_high == TRUE)
]

cantley_filtered <- as_tibble(cantley_filtered_dt) %>%
  filter(!is.na(Kinase), !is.na(Substrate))

# =========================
# STEP 4: Combine with curated data
#  - Pair-level combined set
#  - Site-level combined set
# =========================

# Pair-level from filtered Cantley.
# The exported table now carries the continuous quantities behind every flag
# (motif percentile and rank, promiscuity index, functional score, number of
# shared compartments, and the composite score) so that downstream consumers --
# including the KinaseCanvas Explorer -- can rank and threshold for themselves
# rather than inheriting ours. The Evidence string no longer asserts
# "functional", which was inaccurate for curated sites retained by the
# disjunction above.
cantley_pairs <- cantley_filtered %>%
  transmute(Kinase, Kinase_entrez, Substrate, Substrate_entrez,
            Substrate_Uniprot_ID, Site,
            Curated_site = curated_site,
            Likely_functional = ochoa_high,
            Motif_percentile = Percentile,
            Motif_rank = Rank,
            Promiscuity_index = promiscuity_index,
            Functional_score = functional_score,
            Shared_compartments = n_shared_compartments,
            Motif_evidence = motif_evidence,
            Functional_evidence = functional_evidence,
            Localization_evidence = localization_evidence,
            KinaseCanvas_score = kinasecanvas_score,
            Evidence = "Specificity Atlas (motif-supported, co-localized)",
            Reference_PMID = NA_character_) %>%
  distinct()

# Pair-level from curated (ks_all may have Substrate column named differently; harmonize)
curated_pairs <- kinet_processed %>%
    mutate(residue = str_sub(Site, 1, 1),
         position = suppressWarnings(as.integer(str_sub(Site, 2)))) %>%
  left_join(ochoa_fs_norm, by = c("Substrate_Uniprot_ID", "position")) %>%
  mutate(Likely_functional = !is.na(functional_score) & functional_score >= cutoff) %>%
  mutate(Curated_site = TRUE) %>%
  transmute(Kinase, Kinase_entrez, Substrate, Substrate_entrez,
            Substrate_Uniprot_ID, Site, Curated_site, Likely_functional,
            Motif_percentile = NA_real_, Motif_rank = NA_real_,
            Promiscuity_index = NA_real_,
            Functional_score = functional_score,
            Shared_compartments = NA_integer_,
            Motif_evidence = NA_real_, Functional_evidence = NA_real_,
            Localization_evidence = NA_real_, KinaseCanvas_score = NA_real_,
            Source, Evidence, Reference_PMID) %>%
  distinct()

combined_pairs <- bind_rows(
  cantley_pairs %>% mutate(Source = "Cantley"),
  curated_pairs
) %>%
  distinct()

# ── Flanking sequence (+/-7 residues around the phosphoacceptor) ───────────────
#
# KiNet's own detail panel shows this for curated interactions (Site | Sequence,
# phosphoacceptor bolded), and reviewers repeatedly used flanking-residue context
# (e.g. the -3 position) to argue for or against a specific prediction. We derive
# it once, here, from the canonical UniProt sequence for every site in the
# resource -- curated and predicted alike -- rather than relying on the atlas's
# own SITE_+/-7_AA column, which only covers the Cantley-derived rows. This also
# doubles as a residue-identity check across the *entire* resource, extending
# the same class of verification that caught the KIFAP3 alias-collision error
# to the curated rows, which had never been cross-checked against a canonical
# sequence before.
uniprot_fasta_cache_path <- here("data/uniprot_sequences_cache.fasta")

site_accessions <- combined_pairs %>%
  distinct(Substrate_Uniprot_ID) %>%
  filter(!is.na(Substrate_Uniprot_ID)) %>%
  pull(Substrate_Uniprot_ID)

cached_seqs <- if (file.exists(uniprot_fasta_cache_path)) {
  readAAStringSet(uniprot_fasta_cache_path)
} else {
  AAStringSet()
}
names(cached_seqs) <- sub("\\|.*", "", sub("^[^|]*\\|", "", names(cached_seqs)))  # sp|ACC|NAME -> ACC

missing_accessions <- setdiff(site_accessions, names(cached_seqs))

if (length(missing_accessions) > 0) {
  message("Fetching ", length(missing_accessions), " canonical sequences from UniProt...")
  fetch_batch_size <- 150
  batches <- split(missing_accessions, ceiling(seq_along(missing_accessions) / fetch_batch_size))
  fetched_fasta <- map_chr(batches, function(batch) {
    resp <- request("https://rest.uniprot.org/uniprotkb/accessions") %>%
      req_url_query(accessions = paste(batch, collapse = ","), format = "fasta") %>%
      req_retry(max_tries = 3) %>%
      req_perform()
    resp_body_string(resp)
  })
  # Biostrings' FASTA parser reads from a file path, not a connection.
  tmp_fasta <- tempfile(fileext = ".fasta")
  writeLines(paste(fetched_fasta, collapse = "\n"), tmp_fasta)
  new_seqs <- readAAStringSet(tmp_fasta)
  names(new_seqs) <- sub("\\|.*", "", sub("^[^|]*\\|", "", names(new_seqs)))
  cached_seqs <- c(cached_seqs, new_seqs)
  writeXStringSet(cached_seqs, uniprot_fasta_cache_path)
}

seq_lookup <- as.character(cached_seqs)

flank_radius <- 7L  # residues shown on each side of the phosphoacceptor

extract_flank <- function(acc, residue, position) {
  seq <- seq_lookup[[acc]]
  if (is.null(seq) || is.na(seq) || is.na(position)) return(NA_character_)
  n <- nchar(seq)
  if (position < 1 || position > n) return(NA_character_)
  actual_residue <- substr(seq, position, position)
  if (actual_residue != residue) return(NA_character_)  # verification failed
  start <- max(1, position - flank_radius)
  end   <- min(n, position + flank_radius)
  paste0(
    substr(seq, start, position - 1),
    "[", actual_residue, "]",
    substr(seq, position + 1, end)
  )
}

site_flanks <- combined_pairs %>%
  distinct(Substrate_Uniprot_ID, Site) %>%
  filter(!is.na(Substrate_Uniprot_ID), !is.na(Site)) %>%
  mutate(
    residue  = str_sub(Site, 1, 1),
    position = suppressWarnings(as.integer(str_sub(Site, 2))),
    Flanking_sequence = pmap_chr(
      list(Substrate_Uniprot_ID, residue, position),
      ~ extract_flank(..1, ..2, ..3)
    )
  ) %>%
  select(Substrate_Uniprot_ID, Site, Flanking_sequence)

n_sites_checked <- nrow(site_flanks)
n_sites_verified <- sum(!is.na(site_flanks$Flanking_sequence))
message(sprintf(
  "Flanking sequence / residue-identity check: %d of %d distinct sites verified against canonical UniProt sequences (%.1f%%).",
  n_sites_verified, n_sites_checked, 100 * n_sites_verified / n_sites_checked
))

combined_pairs <- combined_pairs %>%
  left_join(site_flanks, by = c("Substrate_Uniprot_ID", "Site"))

write_csv(combined_pairs, here("results/combined_kinase_substrate_pairs_2025.csv"))

# combined_sites = cantley_filtered %>%
#   mutate(Source = "Cantley") %>%
#   select(Substrate, Substrate_entrez, Site, residue, position, Source) %>%
#   bind_rows(ks_sites_all) %>%
#   distinct()

# write_csv(combined_sites, here("results/combined_kinase_substrate_sites_2025.csv"))

```

# Paper figures

```{r}

combined_pairs = read_csv(here("results/combined_kinase_substrate_pairs_2025.csv")) %>%
  mutate(Source = if_else(Source == "Cantley", "Specificity Atlas", Source))

```

## Sankey plot

```{r}

curated_sources <- c("PhosphoSitePlus","EPSD","iPTMNet","PhosphoELM","PhosphoNetworks")

## 1) Curated substrate universe (for "Novel substrate?" at substrate-level)
curated_sub_universe <- combined_pairs %>%
  filter(Source %in% curated_sources) %>%
  distinct(Substrate) %>%
  pull(Substrate)

## 2) Curated site keys (for "Known site?" at substrate-level)
curated_site_keys <- kinet_processed %>%
  mutate(Site = if ("Site" %in% names(.)) Site else paste0(residue, position)) %>%
  transmute(site_key = make_site_key(Substrate_entrez, Site)) %>%
  distinct() %>% pull(site_key)

## 3) Ochoa functional scores (normalize join keys)
ochoa_fs_norm <- ochoa_fs %>%
  transmute(
    Substrate_Uniprot_ID = norm_uniprot(uniprot),
    position = as.integer(position),
    functional_score
  ) %>%
  distinct()

## 4) Make sure Specificity Atlas universe has clean Site + position + site_key
cantley_uni <- cantley_top %>%
  mutate(
    Site = if ("Site" %in% names(.)) Site else stringr::str_extract(Phosphosite, "^[STY]\\d+$"),
    position = if ("position" %in% names(.)) as.integer(position)
               else suppressWarnings(as.integer(stringr::str_remove(Site, "^[STY]"))),
    site_key = make_site_key(Substrate_entrez, Site)
  ) %>%
  filter(!is.na(Substrate), !is.na(Site), !is.na(position))

## 5) Pair-wise coloc flag, then summarize to substrate-level "any coloc?"
#   (restrict HPA to genes present to speed up)
genes_needed <- unique(c(cantley_uni$Substrate_entrez, cantley_uni$Kinase_entrez))
hpa_entrez = bitr(unique(subcellular_localization$`Gene name`), fromType="SYMBOL", toType="ENTREZID", OrgDb=org.Hs.eg.db)
hpa_long <- subcellular_localization %>%
  select(-Gene) %>%
  rename(Gene = `Gene name`) %>%
  pivot_longer(
    cols = c(contains("Approved"), contains("Supported"), contains("Enhanced")),
    names_to = "tier", values_to = "locs"
  ) %>%
  filter(!is.na(locs)) %>%
  separate_rows(locs, sep = ";") %>%
  mutate(locs = str_trim(locs)) %>%
  filter(locs != "") %>%
  left_join(
    hpa_entrez,
    by = c("Gene" = "SYMBOL")
  ) %>%
  transmute(Gene_entrez = ENTREZID, locs) %>%
  filter(!is.na(Gene_entrez)) %>%   # an NA key would match any gene lacking an Entrez ID
  distinct()

hpa_small <- hpa_long %>%
  filter(Gene_entrez %in% genes_needed) %>%
  distinct(Gene_entrez, locs)

kin_loc <- cantley_uni %>%
  distinct(Kinase, Kinase_entrez) %>%
  mutate(Kinase_entrez = as.character(Kinase_entrez)) %>%
  inner_join(hpa_small, by = c("Kinase_entrez" = "Gene_entrez")) %>%
  transmute(Kinase_entrez, loc = locs) %>%
  filter(!is.na(Kinase_entrez))

sub_loc <- cantley_uni %>%
  distinct(Substrate_entrez) %>%
  mutate(Substrate_entrez = as.character(Substrate_entrez)) %>%
  inner_join(hpa_small, by = c("Substrate_entrez" = "Gene_entrez")) %>%
  transmute(Substrate_entrez, loc = locs) %>%
  filter(!is.na(Substrate_entrez))

coloc_key <- inner_join(kin_loc, sub_loc, by = "loc") %>%
  transmute(Kinase_entrez, Substrate_entrez) %>%
  mutate(
    Kinase_entrez = as.integer(Kinase_entrez),
    Substrate_entrez = as.integer(Substrate_entrez),
    KS_key = paste(Kinase_entrez, Substrate_entrez)
  ) %>%
  distinct()

## 6) Site-level annotations on the Specificity Atlas universe (curated site? Ochoa-high?)
cutoff <- functional_score_cutoff

cantley_sites_annot <- cantley_uni %>%
  # curated site?
  mutate(curated_site = site_key %in% curated_site_keys) %>%
  # join Ochoa FS by UniProt + position
  left_join(
    ochoa_fs_norm,
    by = c("Substrate_Uniprot_ID", "position")
  ) %>%
  mutate(ochoa_high = !is.na(functional_score) & functional_score >= cutoff) %>%
  # The high-confidence filter is applied to the kinase-site pair as a whole: the SAME
  # pair must have a supported site and a co-localized kinase. Evaluating "any supported
  # site" and "any co-localized kinase" separately at the substrate level would keep
  # substrates whose supported sites and co-localized kinases never coincide.
  mutate(KS_key = paste(as.integer(Kinase_entrez), as.integer(Substrate_entrez)),
         coloc = !is.na(fmatch(KS_key, coloc_key$KS_key)),
         triplet_ok = coloc & (curated_site | ochoa_high))

## 7) Collapse to SUBSTRATE-level flags
substrate_summary <- cantley_sites_annot %>%
  group_by(Substrate, Substrate_entrez) %>%
  summarise(
    has_curated_site = any(curated_site, na.rm = TRUE),
    has_functional_site = any(ochoa_high, na.rm = TRUE),
    passes_support = any(curated_site, na.rm = TRUE) | any(ochoa_high, na.rm = TRUE),
    keep = any(triplet_ok, na.rm = TRUE),
    novel_label = if_else(
      any(Substrate %in% curated_sub_universe, na.rm = TRUE),
      "Known substrate",
      "Novel substrate"
    ),
    .groups = "drop"
  ) %>%
  mutate(
    has_curated_site = replace_na(has_curated_site, FALSE),
    has_functional_site = replace_na(has_functional_site, FALSE),
    passes_support = replace_na(passes_support, FALSE),
    keep = replace_na(keep, FALSE)
  )

substrate_paths <- cantley_sites_annot %>%
  distinct(Substrate, Substrate_entrez) %>%
  left_join(substrate_summary, by = c("Substrate", "Substrate_entrez")) %>%
  mutate(
    `Specificity Atlas` = novel_label,
    `Site support` = case_when(
      has_curated_site & has_functional_site ~ "Curated + functional score",
      has_curated_site                      ~ "Curated site",
      has_functional_site                   ~ "Functional score only",
      TRUE                                  ~ "No site-level support"
    ),
    `Co-localized?` = case_when(
      !passes_support ~ NA_character_,
      keep ~ "Co-localized",
      TRUE ~ "No co-localization"
    ),
    `Filtered Result` = if_else(keep, novel_label, NA_character_)
  )

stage_vars <- c("Specificity Atlas", "Site support", "Co-localized?", "Filtered Result")

# Per-stage counts for the alluvial, so readers can see how many substrates enter
# and leave each stratum rather than having to read them off the ribbons.
alluvial_stage_counts <- substrate_paths %>%
  select(all_of(stage_vars)) %>%
  pivot_longer(everything(), names_to = "stage", values_to = "stratum") %>%
  filter(!is.na(stratum)) %>%
  count(stage, stratum, name = "n_substrates") %>%
  mutate(stage = factor(stage, levels = stage_vars)) %>%
  arrange(stage, desc(n_substrates))

# The last stage must be the substrate set of the released atlas layer (Figure 1B); if this
# fails, the alluvial is no longer describing the filter the pipeline actually applies.
stopifnot(
  sum(alluvial_stage_counts$n_substrates[alluvial_stage_counts$stage == "Filtered Result"]) ==
    dplyr::n_distinct(combined_pairs$Substrate[combined_pairs$Source == "Specificity Atlas"])
)

write_csv(alluvial_stage_counts, here("results/alluvial_stage_counts.csv"))
print(as.data.frame(alluvial_stage_counts), row.names = FALSE)

alluv_df <- substrate_paths %>%
  select(all_of(stage_vars)) %>%
  group_by(across(everything())) %>%
  summarise(n = n(), .groups = "drop") %>%
  arrange(desc(n))

alluv_long <- alluv_df %>%
  mutate(id = row_number()) %>%
  tidyr::pivot_longer(cols = all_of(stage_vars), names_to = "stage", values_to = "label") %>%
  mutate(stage = factor(stage, levels = stage_vars)) %>%
  filter(!is.na(label))

## 9) Plot (substrate-level Sankey)
pal_stage <- c(
  "Specificity Atlas substrates"   = "#2C7BB6",
  "Curated + functional score"     = "#00A6D6",
  "Curated site"                   = "#3E8FA8",
  "Functional score only"          = "#7FCDE0",
  "No site-level support"          = "#BFC0C0",
  "Co-localized"                   = "#7BC950",
  "No co-localization"             = "#F25F5C",
  "Known substrate"                = "#4355B9",
  "Novel substrate"                = "#F9A03F"
)
# Look each label up by name and fall back to grey, rather than filtering the
# palette and renaming positionally -- the latter silently mismatches (and then
# errors) as soon as a new stratum label is introduced.
lab_levels <- unique(alluv_long$label)
pal_use <- setNames(
  unname(ifelse(lab_levels %in% names(pal_stage), pal_stage[lab_levels], "#B0B0B0")),
  lab_levels
)

p_sankey_substrates <- ggplot(
  alluv_long,
  aes(x = stage, stratum = label, alluvium = id, y = n,
      fill = label, label = label)
) +
  geom_flow() +
  geom_stratum() +
  geom_text(stat = "stratum", size = 3.2, color = "black") +
  scale_fill_manual(values = pal_use, guide = "none") +
  scale_y_continuous(expand = expansion(mult = c(0.02, 0.08)), labels = scales::comma) +
  labs(
    x = NULL, y = "Substrate count", fill = NULL) +
  theme_pubr()

p_sankey_substrates

```

## Figure 1
```{r}

dataset_label_colors <- c(
  "Specificity Atlas" = "#B279A2",
  "PhosphoSitePlus" = "#4C78A8",
  "EPSD" = "#F58518",
  "iPTMNet" = "#E45756",
  "Total" = "#000000"
)

color_dataset_labels <- function(labels) {
  map_chr(labels, function(lbl) {
    out <- lbl
    for (nm in names(dataset_label_colors)) {
      out <- str_replace_all(
        out,
        fixed(nm),
        glue("<span style='color:{dataset_label_colors[[nm]]}'>{nm}</span>")
      )
    }
    out
  })
}

p1A <- p_sankey_substrates +
  theme(legend.position = "none")+
  labs(tag="C")

fig1B_df <- combined_pairs %>%
  group_by(Source) %>%
  summarise(
    Kinases    = n_distinct(Kinase),
    Substrates = n_distinct(Substrate)
  ) %>%
  ungroup() %>%
  #Add in total numbers
  bind_rows(tibble(
    Source = "Total",
    Kinases = n_distinct(combined_pairs$Kinase),
    Substrates = n_distinct(combined_pairs$Substrate)
  )) %>%
   tidyr::pivot_longer(-Source, names_to="Entity", values_to="Count")

p1B <- ggplot(fig1B_df, aes(Source, Count, fill=Source))+
  geom_col(width=0.65, color="grey20", show.legend = FALSE)+
  scale_fill_manual(values=c(
    "PhosphoSitePlus"="#4C78A8",
    "EPSD"="#F58518",
    "iPTMNet"="#E45756",
    "Specificity Atlas"="#72B7B2",
    "Total"="#54A24B"
  ))+
  scale_x_discrete(labels = color_dataset_labels)+
  scale_y_continuous(labels=scales::comma, expand=expansion(mult=c(0,0.1)))+ # 10% headroom
  geom_text(aes(label=Count), vjust=-0.3, size=3.2)+
  facet_wrap(~Entity, scales="free_y")+
  theme_pubr()+
  theme(strip.background = element_rect(fill="grey90", color="black"),
        strip.text = element_text(face="bold"),
        axis.text.x = ggtext::element_markdown(angle = 30, hjust = 1))+
  labs(x=NULL, y="Count", tag="B")

deg_kin <- combined_pairs %>%
  filter(Source %in% c("Specificity Atlas","PhosphoSitePlus","EPSD","iPTMNet")) %>%
  count(Kinase, name="n_substrates")

p1C <- ggplot(deg_kin, aes(n_substrates))+
  geom_histogram(bins=60, fill="#4C78A8", color="grey20")+
  geom_vline(aes(xintercept = median(n_substrates, na.rm=TRUE)),
             linetype="dashed", color="black")+
  annotate("text", x=median(deg_kin$n_substrates), y=Inf,
           label=paste0("Median = ", median(deg_kin$n_substrates)), vjust=1.5, hjust=-0.1, size=3.2)+
  scale_x_continuous(trans="log1p")+
  scale_y_continuous(labels=comma, expand=expansion(mult=c(0,0.1)))+ # 10% headroom
  theme_pubr() +
  labs(title="Substrates per kinase (final resource)", x="Substrates per kinase (log1p)", y="Count", tag="D")

# Build per-pair set membership as list-column for ggupset
pairs_src_long <- combined_pairs %>%
  distinct() %>%
  select(Kinase, Substrate, Source) %>%
  distinct()

pairs_set <- pairs_src_long %>%
  group_by(Substrate) %>%
  summarise(Sets = list(sort(unique(Source))), .groups="drop")

p1D <- ggplot(pairs_set, aes(x = Sets))+
  geom_bar(fill="#4C78A8", color="grey20")+
  scale_x_upset(order_by = "freq", n_intersections = 20, sets = c("PhosphoSitePlus","EPSD","iPTMNet","Specificity Atlas"))+
  #geom_text(stat='count', aes(label=after_stat(count)), vjust=-0.3, size=3.2)+
  scale_y_continuous(labels=comma, expand=expansion(mult=c(0,0.1)))+ # 10% headroom
  labs(y="# of Substrates", x=NULL, tag ="D")+
  theme_pubr()

#Number kinases and substrates per family in the final set
p1E_df <- combined_pairs %>%
  left_join(kinase_list, by="Kinase") %>%
  group_by(Family) %>%
  summarise(
    Kinases    = n_distinct(Kinase),
    Substrates = n_distinct(Substrate)
  ) %>%
  ungroup() %>%
  tidyr::pivot_longer(-Family, names_to="Entity", values_to="Count")

p1E <- p1E_df %>%
  filter(!is.na(Family)) %>%
  ggplot(aes(reorder_within(Family, desc(Count), Entity), Count, fill=Entity))+
  scale_fill_manual(values=c("Kinases"="#4C78A8", "Substrates"="#F58518"))+
  facet_wrap(~Entity, scales="free")+
  geom_col(width=0.65, color="grey20", show.legend=FALSE)+
  #geom_text(aes(label=Count), vjust=-0.3, size=3.2)+
  scale_x_reordered()+
  #scale_y_continuous(labels=comma, expand=expansion(mult=c(0,0.1)))+ # 10% headroom
    theme_pubr(x.text.angle = 45)+
  theme(strip.background = element_rect(fill="grey90", color="black"),
        strip.text = element_text(face="bold"))+
  labs(x=NULL, y="Count", tag="E")

# Load BioRender figure for Panel A
# path: figures/paper_figures/KS resource figure.png
biorender_img <- image_read(here("figures/paper_figures/KS resource figure.png"))
p1_biorender <- ggplot() + 
  draw_image(biorender_img) + 
  theme_void() + 
  labs(tag = "A")

# First row: BioRender (Panel A) and Source counts (Panel B)
row1 <- p1_biorender

row2 <- p1B

# Second row: Alluvial (Panel C)
row3 <- p1A

# Third row: UpSet (Panel D) and Family (Panel E)
row4 <- p1D + p1E + plot_layout(widths = c(1, 2))

# Combine with relative heights
# BioRender image is 2038x688 (approx 3:1 width:height)
# We adjust the relative height of row1 to 0.5 to keep it from taking too much vertical space,
# and increase the ggsave height to compensate.
fig1 <- (row1 / row2 / row3 / row4) +
  plot_layout(heights = c(1.2, 1, 1.4, 0.8)) &
  theme(plot.tag = element_text(face = "bold"))
fig1
# Original width ~10in, height ~10.2in. 
# Adding Panel A requires more vertical space. Increasing height to 12.5in.
ggsave(here("figures/paper_figures/fig1.png"), fig1, width=10, height=14, dpi=500)
```

## Figure S1
```{r}

figS1A_df <- combined_pairs %>%
  group_by(Source) %>%
  summarise(
    Substrates = n_distinct(Substrate),
    Sites      = n_distinct(paste(Substrate, Site))
  ) %>% 
  ungroup() %>%
  #Add in total numbers
  bind_rows(tibble(
    Source = "Total",
    Substrates = n_distinct(combined_pairs$Substrate),
    Sites = n_distinct(paste(combined_pairs$Substrate, combined_pairs$Site))
  )) %>%
  tidyr::pivot_longer(-Source, names_to="Entity", values_to="Count")

pS1A <- ggplot(figS1A_df %>% filter(Entity == "Sites"), aes(Source, Count, fill=Source))+
  geom_col(width=0.65, color="grey20", show.legend=FALSE)+
  scale_fill_manual(values=c(
    "PhosphoSitePlus"="#4C78A8",
    "EPSD" ="#F58518",
    "iPTMNet"="#E45756",
    "Specificity Atlas"="#72B7B2",
    "Total"="#54A24B"
  ))+
  geom_text(aes(label=Count), vjust=-0.3, size=3.2)+
  facet_wrap(~Entity, scales="free_y")+
  theme_pubr(x.text.angle = 45)+
  theme(strip.background = element_rect(fill="grey90", color="black"),
        strip.text = element_text(face="bold"))+
  scale_y_continuous(labels=comma, expand=expansion(mult=c(0,0.1)))+ # 10% headroom
  labs(x=NULL, y="Count")

# Build per-pair set membership as list-column for ggupset
sites_src_long <- combined_pairs %>%
  distinct() %>%
  select(Substrate, Site, Source) %>%
  distinct()

sites_set <- sites_src_long %>%
  group_by(Substrate, Site) %>%
  summarise(Sets = list(sort(unique(Source))), .groups="drop")

pS1B <- ggplot(sites_set, aes(x = Sets))+
  geom_bar(fill="#4C78A8", color="grey20")+
  scale_x_upset(order_by = "freq", n_intersections = 20, sets = c("PhosphoSitePlus","EPSD","iPTMNet","Specificity Atlas"))+
  scale_y_continuous(labels=comma, expand=expansion(mult=c(0,0.1)))+ # 10% headroom
  # geom_text(stat='count', aes(label=after_stat(count)), vjust=-0.3, size=3.2)+
  labs(y="Site count", x=NULL)+
  theme_pubr()

curated_sources <- c("PhosphoSitePlus","EPSD","iPTMNet")

# 1) Start from site-by-source table; add UniProt if missing by joining ks_sites_all
sites_src_long2 <- combined_pairs %>%
  distinct(Substrate, Site, Source) %>%
  # try to bring UniProt & Entrez from your site union table if not present
  left_join(
    kinet_processed %>%
      distinct(Substrate, Site, Substrate_Uniprot_ID, Substrate_entrez),
    by = c("Substrate","Site")
  ) %>%
  mutate(
    Substrate_Uniprot_ID = norm_uniprot(Substrate_Uniprot_ID),
    position = suppressWarnings(as.integer(str_remove(Site, "^[STY]")))
  )

# 2) Label whether a *site (union across sources)* is curated (present in *any* curated source)
site_curated_union <- sites_src_long2 %>%
  mutate(is_cur_src = Source %in% curated_sources) %>%
  group_by(Substrate, Site) %>%
  summarise(curated_union = any(is_cur_src), .groups = "drop")

sites_src_annot <- sites_src_long2 %>%
  left_join(site_curated_union, by = c("Substrate","Site"))

# 3) Join Ochoa functional scores (UniProt + position), mark functional threshold (>0.5)
ochoa_fs_norm <- ochoa_fs %>%
  transmute(Substrate_Uniprot_ID = norm_uniprot(uniprot),
            position = as.integer(position),
            functional_score)

sites_src_annot <- sites_src_annot %>%
  left_join(ochoa_fs_norm, by = c("Substrate_Uniprot_ID","position")) %>%
  filter(!is.na(functional_score)) %>%
  mutate(functional_high = functional_score >= cutoff)

s1c_pct <- sites_src_annot %>%
  group_by(Source) %>%
  summarise(
    n_sites_src = n(),                                       # sites in this source
    n_func_src  = sum(functional_high, na.rm = TRUE),        # functional within this source
    pct_func    = n_func_src / n_sites_src                   # fraction
  ) %>%
  ungroup()

# also add a "Total" bar across the union of sites (not double-counting across sources)
total_union <- sites_src_annot %>%
  distinct(Substrate, Site, functional_high) %>%
  summarise(
    Source      = "Total",
    n_sites_src = n(),
    n_func_src  = sum(functional_high, na.rm = TRUE),
    pct_func    = n_func_src / n_sites_src
  )

s1c_pct <- bind_rows(
  s1c_pct,
  total_union
) %>%
  arrange(Source)

# ---- 5) Plot: single bar = % functional sites --------------------------------
pS1C <- ggplot(s1c_pct, aes(Source, pct_func, fill = Source)) +
  geom_col(width = 0.65, color = "grey20", show.legend = FALSE) +
  geom_text(aes(label = scales::percent(pct_func, accuracy = 0.1)),
            vjust = -0.3, size = 3.2) +
  scale_fill_manual(values = c(
    "PhosphoSitePlus" = "#4C78A8",
    "EPSD"            = "#F58518",
    "iPTMNet"         = "#E45756",
    "Specificity Atlas" = "#72B7B2",
    "Total"           = "#54A24B"
  )) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1),
                     expand  = expansion(mult = c(0, 0.1))) +  # 10% headroom
  labs(x = NULL, y = "% Sites Functional") +
  theme_pubr(x.text.angle = 45) +
  theme(strip.background = element_rect(fill = "grey90", color = "black"),
        strip.text       = element_text(face = "bold"))


figS1 <- (pS1A | pS1B) / pS1C  +
  plot_annotation(tag_levels = 'A') &
  theme(plot.tag = element_text(face = "bold"))
figS1
ggsave(here("figures/paper_figures/figS1.png"), figS1, width=9, height=7, dpi=600)

```

## Figure 2
```{r}

# ── A1. Cantley at the motif percentile cutoff, site level (no coloc yet) ─────────────────────────
cantley_sites90 <- cantley_top

# ── A2. Curated-site flag (PSP/ELM/PN union) ──────────────────────────────────
curated_keys <- kinet_processed %>%
  mutate(Site = if ("Site" %in% names(.)) Site else paste0(residue, position)) %>%
  transmute(key = paste(Substrate_entrez, Site, sep=":")) %>%
  distinct()

cantley_sites90 <- cantley_sites90 %>%
  mutate(key = paste(Substrate_entrez, Site, sep=":")) %>%
  mutate(curated_site = key %in% curated_keys$key)

# ── A3. Ochoa functional score mapping & flag (>0.5) ──────────────────────────
ochoa_fs_norm <- ochoa_fs %>%
  transmute(Substrate_Uniprot_ID = norm_uniprot(uniprot),
            position = as.integer(position),
            functional_score)

cantley_sites90 <- cantley_sites90 %>%
  left_join(ochoa_fs_norm, by = c("Substrate_Uniprot_ID","position")) %>%
  mutate(ochoa_high = !is.na(functional_score) & functional_score >= cutoff)

# ── A4. NEW site filter: keep if curated OR Ochoa-high ────────────────────────
cantley_sites90_keep <- cantley_sites90 %>%
  filter(curated_site | ochoa_high)

# ── A5. Derive KS pairs from the kept sites (gene-level) ──────────────────────
cantley_pairs90_keep <- cantley_sites90_keep %>%
  distinct(Kinase, Kinase_entrez, Substrate, Substrate_entrez)

# ── B1. HPA subset to relevant genes (kept set only) ──────────────────────────
genes_needed <- unique(c(cantley_pairs90_keep$Kinase_entrez,
                         cantley_pairs90_keep$Substrate_entrez))

hpa_small <- hpa_long %>%
  filter(Gene_entrez %in% genes_needed) %>%
  distinct(Gene_entrez, locs)

kin_loc <- hpa_small %>% rename(Kinase_entrez = Gene_entrez, loc = locs)
sub_loc <- hpa_small %>% rename(Substrate_entrez = Gene_entrez, loc = locs)

# ── B2. Per-kinase denominators (all kept vs HPA-annotatable kept) ────────────
annot_subs <- unique(sub_loc$Substrate_entrez)

denoms <- cantley_pairs90_keep %>%
  group_by(Kinase, Kinase_entrez) %>%
  summarise(
    n_all_subs   = n_distinct(Substrate_entrez),
    n_annot_subs = n_distinct(if_else(Substrate_entrez %in% annot_subs,
                                      Substrate_entrez, NA_character_)),
    .groups = "drop"
  )

kin_with_hpa <- unique(kin_loc$Kinase_entrez)

denoms <- denoms %>%
  mutate(kin_has_hpa = Kinase_entrez %in% kin_with_hpa) %>%
  mutate(n_annot_subs = ifelse(kin_has_hpa, n_annot_subs, 0L))

# ── B3. Count co-localized substrates per kinase (≥1 shared compartment) ──────
ks_coloc_rows <- cantley_pairs90_keep %>%
  inner_join(kin_loc, by = "Kinase_entrez") %>%
  inner_join(sub_loc, by = c("Substrate_entrez","loc")) %>%
  distinct(Kinase_entrez, Substrate_entrez)

coloc_counts <- ks_coloc_rows %>%
  group_by(Kinase_entrez) %>%
  summarise(n_coloc_subs = n_distinct(Substrate_entrez), .groups="drop")

# ── B4. Merge & compute fractions; build plots ────────────────────────────────
per_kinase_coloc <- denoms %>%
  left_join(coloc_counts, by = "Kinase_entrez") %>%
  mutate(
    n_coloc_subs   = replace_na(n_coloc_subs, 0L),
    pct_coloc_all  = ifelse(n_all_subs   > 0, 100 * n_coloc_subs / n_all_subs, NA_real_),
    pct_coloc_annot= ifelse(kin_has_hpa & n_annot_subs > 0, 100 * n_coloc_subs / n_annot_subs, NA_real_)
  ) %>%
  arrange(desc(pct_coloc_annot))

# A percentage cannot exceed 100. This guard exists because NA Entrez IDs once
# made a numerator pool across several kinases and pushed values past 6000%.
stopifnot(max(per_kinase_coloc$pct_coloc_annot, na.rm = TRUE) <= 100,
          max(per_kinase_coloc$pct_coloc_all,   na.rm = TRUE) <= 100)

# Histogram + density + median (same look)
p2B_dist <- ggplot(per_kinase_coloc, aes(pct_coloc_all)) +
  geom_histogram(bins = 40, fill = "#4C78A8", color = "grey20", alpha = 0.9) +
  geom_density(aes(y = after_stat(..count..)), linewidth = 0.8, color = "#2F4B7C", alpha = 0.6) +
  geom_vline(xintercept = median(per_kinase_coloc$pct_coloc_all, na.rm = TRUE),
             linetype = "dashed", color = "black") +
  annotate("text",
           x = median(per_kinase_coloc$pct_coloc_all, na.rm = TRUE),
           y = Inf,
           label = paste0("Median = ", round(median(per_kinase_coloc$pct_coloc_all, na.rm=TRUE), 1), "%"),
           vjust = 2.5, hjust = -0.05, size = 3.1) +
  scale_x_continuous(labels = function(x) paste0(x, "%"), expand = expansion(mult = c(0, 0.02))) +
  scale_y_continuous(labels = scales::comma, expand = expansion(mult = c(0, 0.1))) +  # 10% headroom
  labs(x = "% Predicted Substrates\nCo-localized per Kinase",
       y = "Count") +
  theme_pubr()

# k_top <- per_kinase_coloc %>% slice_max(pct_coloc_annot, n = 15, with_ties = FALSE)
# k_bot <- per_kinase_coloc %>% filter(!is.na(pct_coloc_annot)) %>% slice_min(pct_coloc_annot, n = 15, with_ties = FALSE)

# p_top <- ggplot(k_top, aes(x = pct_coloc_annot, y = fct_reorder(Kinase, pct_coloc_annot))) +
#   geom_col(fill = "#54A24B", color = "grey20") +
#   geom_text(aes(label = sprintf("%.1f%%", round(pct_coloc_annot, 1))), hjust = -0.1, size = 3.0) +
#   scale_x_continuous(labels = function(x) paste0(x, "%"), expand = expansion(mult = c(0, 0.4))) +
#   labs(x = "% substrates co-localized (annotatable)", y = NULL) +
#   theme_pubr() +
#   theme(legend.position = "none") +
#   coord_cartesian(clip = "off")

# p_bot <- ggplot(k_bot, aes(x = pct_coloc_annot, y = fct_reorder(Kinase, pct_coloc_annot))) +
#   geom_col(fill = "#E45756", color = "grey20") +
#   geom_text(aes(label = sprintf("%.1f%%", round(pct_coloc_annot, 1))), hjust = -0.1, size = 3.0) +
#   scale_x_continuous(labels = function(x) paste0(x, "%"), expand = expansion(mult = c(0, 0.08))) +
#   labs(title = "Bottom kinases by co-localized substrates",
#        x = "% co-localized (annotatable)", y = NULL) +
#   theme_pubr() +
#   coord_cartesian(clip = "off")

# One observation per unique site: a site's support does not depend on which kinase
# matched it, so averaging over kinase-site pairs would weight each site by the number
# of kinases whose motif it fits. This is also the unit of the chi-square test below.
p2C <- cantley_sites90 %>%
  distinct(Substrate_Uniprot_ID, Site, curated_site, ochoa_high) %>%
  mutate(pass = curated_site | ochoa_high,
         class = case_when(substr(Site,1,1) == "S" ~ "Serine (S)",
                           substr(Site,1,1) == "T" ~ "Threonine (T)",
                           substr(Site,1,1) == "Y" ~ "Tyrosine (Y)",
                           TRUE ~ "Other")) %>%
  filter(class != "Other") %>%
  group_by(class) %>%
  summarise(frac = mean(pass, na.rm = TRUE), .groups = "drop") %>%
  mutate(class = factor(class, levels = c("Serine (S)","Threonine (T)","Tyrosine (Y)"))) %>%
  ggplot(aes(class, frac, fill = class)) +
  geom_col(width = 0.7, color = "grey20") +
  geom_text(aes(label = sprintf("%.1f%%", frac*100)), vjust = -0.3, size = 3.2) +
  scale_fill_manual(values = c("Serine (S)"="#B279A2", "Threonine (T)"="#B279A2", "Tyrosine (Y)"="#4C78A8")) +
  scale_y_continuous(labels = percent, expand = expansion(mult = c(0, 0.1))) +
  labs(x = "Residue class", y = "% Sites Passing") +
  theme_pubr(x.text.angle = 45) +
  theme(legend.position = "none")

# row1 <- p2B_dist + p_top + plot_layout(widths = c(2, 1))
# row2 <- p2C + p2D + plot_layout(widths = c(1, 2))

fig2 <- p2B_dist + (p2C + pS1C) +
  plot_annotation(tag_levels = 'A') &
  theme(plot.tag = element_text(face = "bold"))

fig2
ggsave(here("figures/paper_figures/fig2.png"), fig2, width=10, height=5, dpi=400)
```

## Figure 3

```{r}
pal <- list(
  curated = "#4C78A8",
  cantley = "#72B7B2",
  novel   = "#54A24B",
  line    = "grey20"
)

curated_sources <- c("PhosphoSitePlus","EPSD","iPTMNet")

# 0) Identify overlapping kinases
kinase_by_src <- combined_pairs %>%
  distinct(Kinase, Source)

kins_spec_atlas <- kinase_by_src %>% filter(Source == "Specificity Atlas") %>% pull(Kinase) %>% unique()
kins_curated <- kinase_by_src %>% filter(Source %in% curated_sources) %>% pull(Kinase) %>% unique()

overlap_kins <- intersect(kins_spec_atlas, kins_curated)

# 1) Build pair-level flags within the overlap
pairs_flags_overlap <- combined_pairs %>%
  #filter(Kinase %in% overlap_kins) %>%
  distinct(Kinase, Substrate, Source) %>%
  group_by(Kinase, Substrate) %>%
  summarise(
    is_curated = any(Source %in% curated_sources),
    in_spec_atlas = any(Source == "Specificity Atlas"),
    .groups = "drop"
  )

# ── 3A. New pairs added by Specificity Atlas (restricted to overlap kinases) ────────────
# pair-level flags already built in `pairs_flags_overlap`
cov_tbl <- pairs_flags_overlap %>%
  summarise(
    curated_pairs   = sum(is_curated),                 # existing curated pairs
    new_pairs_added = sum(in_spec_atlas & !is_curated)    # Specificity Atlas-only pairs
  ) %>%
  tidyr::pivot_longer(everything(), names_to = "metric", values_to = "count") %>%
  mutate(metric = factor(metric,
                         levels = c("curated_pairs","new_pairs_added"),
                         labels = c("Curated pairs","Novel pairs")))

p3A_coverage <- ggplot(cov_tbl, aes(x = metric, y = count, fill = metric)) +
  geom_col(width = 0.65, color = pal$line, show.legend = FALSE) +
  geom_text(aes(label = scales::comma(count)), vjust = -0.3, size = 3.2) +
  scale_fill_manual(values = c("Curated pairs" = pal$curated,
                               "Novel pairs" = pal$novel)) +
  scale_y_continuous(labels = scales::comma, expand = expansion(mult = c(0, 0.12))) +
  annotate(
    "text", x = 1.5, y = max(cov_tbl$count) * 1.2,
    label = {
      df <- tidyr::pivot_wider(cov_tbl, names_from = metric, values_from = count)
      # increase relative to the curated set, i.e. novel / curated (the same
      # quantity as the fold-change quoted in the text: total / curated = 1 + this)
      inc <- df$`Novel pairs` / df$`Curated pairs`
      paste0("Increase = +", scales::percent(inc, accuracy = 0.1))
    },
    size = 3.3, fontface = 2
  ) +
  labs(x = NULL, y = "Pair count") +
  theme_pubr()

# ── 3B. Distribution of novel substrates per kinase (overlap kinases only) ────
curated_vs_novel_edges_overlap <- combined_pairs %>%
  filter(Kinase %in% overlap_kins) %>%
  distinct(Kinase, Substrate, Source) %>%
  mutate(is_curated = Source %in% curated_sources,
         is_spec_atlas = Source == "Specificity Atlas") %>%
  group_by(Kinase, Substrate) %>%
  summarise(
    known = any(is_curated),
    # pair-level: predicted and never curated (a per-row test counted the 881
    # atlas pairs that are also curated as novel)
    novel = any(is_spec_atlas) & !any(is_curated),
    .groups = "drop"
  ) %>%
  group_by(Kinase) %>%
  summarise(
    n_known = sum(known),
    n_novel = sum(novel),
    n_total = n_known + n_novel,
    frac_known = ifelse(n_total > 0, n_known / n_total, NA_real_),
    .groups = "drop"
  )

p3B <- ggplot(curated_vs_novel_edges_overlap, aes(x = n_novel)) +
  geom_histogram(fill = pal$novel, color = pal$line, bins = 40) +
  #add line for median
  geom_vline(aes(xintercept = median(n_novel, na.rm=TRUE)),
             linetype = "dashed", color = "black") +
  annotate("text",
           x = median(curated_vs_novel_edges_overlap$n_novel, na.rm=TRUE),
           y = Inf,
           label = paste0("Median = ", median(curated_vs_novel_edges_overlap$n_novel, na.rm=TRUE)),
           vjust = 1.5, hjust = -0.1, size = 3.2) +
  theme_pubr() +
  labs(x = "# Novel substrates per kinase", y = "Kinase count")

# ── 3C. Kinase-wise expansion (gain) in overlap kinases ───────────────────────
kin_expansion_overlap <- combined_pairs %>%
  filter(Kinase %in% overlap_kins) %>%
  distinct(Kinase, Substrate, Source) %>%
  group_by(Kinase) %>%
  summarise(
    n_known = n_distinct(Substrate[Source %in% curated_sources]),
    n_novel = n_distinct(Substrate[Source == "Specificity Atlas" &
                                     !(Substrate %in% Substrate[Source %in% curated_sources])]),
    n_total = n_known + n_novel,
    gain    = ifelse(n_known > 0, n_total / n_known, Inf),
    .groups = "drop"
  )

top_gain <- kin_expansion_overlap %>%
  arrange(desc(gain)) %>%
  slice_head(n = 20)

p3C <- ggplot(top_gain, aes(x = gain, y = fct_reorder(Kinase, gain))) +
  geom_col(fill = pal$cantley, color = pal$line) +
  geom_text(aes(label = paste0("+", n_novel)), hjust = -0.1, size = 3) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.2))) +
  labs(x = "Expansion factor (total / curated substrates)", y = NULL,
       title = "Kinase-wise expansion from Specificity Atlas (same kinases)") +
  theme_pubr()

# ── 3D. Family-wise novelty; show only families with >15 dark kinases ─────────
understudied_fams <- all_kinases %>%
  filter(class == "Dark") %>%
  left_join(kinase_list, by = c("symbol" = "Kinase")) %>%
  filter(!is.na(Family)) %>%
  count(Family, name = "n_dark") %>%
  filter(n_dark > 15) %>%                    # <- keep only >15
  arrange(desc(n_dark))

kin_fam_novelty_overlap <- combined_pairs %>%
  #filter(Kinase %in% overlap_kins) %>%
  distinct(Kinase, Substrate, Source) %>%
  left_join(kinase_list, by = "Kinase") %>%
  filter(!is.na(Family)) %>%
  group_by(Family) %>%
  summarise(
    n_known = n_distinct(Substrate[Source %in% curated_sources]),
    n_novel = n_distinct(Substrate[Source == "Specificity Atlas" &
                         !(Substrate %in% Substrate[Source %in% curated_sources])]),
    frac_novel = ifelse(n_known + n_novel > 0, n_novel / (n_known + n_novel), NA_real_),
    .groups = "drop"
  ) %>%
  inner_join(understudied_fams, by = "Family") %>%   # <- restrict to >15 dark families
  arrange(desc(frac_novel))

p3D <- ggplot(kin_fam_novelty_overlap,
              aes(x = frac_novel, y = fct_reorder(Family, frac_novel))) +
  geom_col(aes(fill = n_dark), color = pal$line) +
  scale_fill_gradient(low = "#4C78A8", high = "#B279A2", name = "# Understudied Kinases") +
  geom_text(aes(label = scales::percent(frac_novel, accuracy = 0.1)), hjust = -0.1, size = 3) +
  scale_x_continuous(labels = scales::percent, expand = expansion(mult = c(0, 0.15))) +
  labs(x = "% Novel Substrates", y = "Kinase Family") +
  theme_pubr()

# curated sources
curated_sources <- c("PhosphoSitePlus","EPSD","iPTMNet","PhosphoELM","PhosphoNetworks")

# ── 3E. % of pairs involving Understudied (Dark) kinases: Curated vs Curated+Specificity Atlas ──
kin_class <- all_kinases %>%
  transmute(Kinase_entrez = as.character(entrez_id),
            Kinase_class  = if_else(class_2019 == "Dark", "Understudied (Dark)", "Well-studied (Light)"))

# kinase_list = read_xlsx(here("data/Comprehensive_Kinase_Simple.xlsx")) %>%
#   mutate(Kinase_entrez = as.character(`Human Gene ID`)) %>%
#   left_join(kin_class, by = "Kinase_entrez") %>%
#   write_csv(here("data/Comprehensive_Kinase_Simple_with_class.csv"))

curated_only_pairs <- combined_pairs %>%
  filter(Source %in% curated_sources) %>%
  distinct(Kinase, Kinase_entrez, Substrate, Substrate_entrez) %>%
  mutate(Set = "Curated")

proposed_pairs <- combined_pairs %>%
  filter(Source %in% c(curated_sources, "Specificity Atlas")) %>%
  distinct(Kinase, Kinase_entrez, Substrate, Substrate_entrez) %>%
  mutate(Set = "Curated + Specificity")

pairs_for_compare <- bind_rows(curated_only_pairs, proposed_pairs) %>%
  mutate(Kinase_entrez = as.character(Kinase_entrez)) %>%
  left_join(kin_class, by = "Kinase_entrez") %>%
  filter(!is.na(Kinase_class)) %>%
  mutate(Set = factor(Set, levels = c("Curated","Curated + Specificity")))

dark_fracs <- pairs_for_compare %>%
  mutate(is_dark = Kinase_class == "Understudied (Dark)") %>%
  group_by(Set) %>%
  summarise(frac_dark = mean(is_dark), n = n(), .groups = "drop")

p3E_pairs <- ggplot(dark_fracs, aes(x = Set, y = frac_dark)) +
  geom_col(color = "grey20", fill = "#B279A2", width = 0.7, show.legend = FALSE) +
  geom_text(aes(label = scales::percent(frac_dark, accuracy = 0.1)),
            vjust = -0.3, size = 3.1) +
  scale_y_continuous(labels = scales::percent, limits = c(0, NA),
                     expand = expansion(mult = c(0, 0.1))) +
  labs(x = NULL, y = "% of KS pairs with\nDark Kinases") +
  theme_pubr()

fig3 <- (p3D | p3E_pairs) / (p3A_coverage | p3B) +
  plot_annotation(tag_levels = 'A') &
  theme(plot.tag = element_text(face = "bold"))
fig3
ggsave(here("figures/paper_figures/fig3.png"), fig3, width=9, height=7, dpi=600)

```

## Manuscript numbers

Every statistic quoted in the manuscript is computed here and written to a single
file, so that the text and the pipeline cannot drift apart between revisions.

```{r}
curated_sources_final <- c("PhosphoSitePlus", "EPSD", "iPTMNet")
# combined_pairs is re-read and relabelled upstream, so atlas rows carry the
# label "Specificity Atlas" rather than "Cantley" by the time we get here.
atlas_source_label <- if ("Specificity Atlas" %in% combined_pairs$Source)
  "Specificity Atlas" else "Cantley"

pairs_curated <- combined_pairs %>%
  filter(Source %in% curated_sources_final) %>%
  distinct(Kinase, Substrate)
pairs_atlas <- combined_pairs %>%
  filter(Source == atlas_source_label) %>%
  distinct(Kinase, Substrate)
pairs_novel <- anti_join(pairs_atlas, pairs_curated, by = c("Kinase", "Substrate"))

novel_per_kinase <- pairs_novel %>% count(Kinase, name = "n_novel")

# Matches the definition plotted in Figure 2B: one row per unique motif-passing
# site, "pass" = curated OR functional-score support. `pct_func_only`
# is the functional-score-only figure, restricted to sites that actually carry a
# score, which is the quantity the running text refers to.
residue_pass <- cantley_sites90 %>%
  mutate(residue = substr(Site, 1, 1)) %>%
  filter(residue %in% c("S", "T", "Y")) %>%
  distinct(Substrate_Uniprot_ID, Site, residue, curated_site, ochoa_high, functional_score) %>%
  group_by(residue) %>%
  summarise(
    n            = n(),
    n_sites      = n_distinct(paste(Substrate_Uniprot_ID, Site)),
    pct_pass     = 100 * mean(curated_site | ochoa_high, na.rm = TRUE),
    pct_func_only = 100 * mean(ochoa_high[!is.na(functional_score)], na.rm = TRUE),
    .groups = "drop"
  )

# Classify by Entrez ID, exactly as Figure 3B does: gene symbols are not a stable
# key here (DarkKinaseTools still uses superseded symbols for a few genes).
kin_class_final <- all_kinases %>%
  transmute(Kinase_entrez = as.character(entrez_id),
            Kinase_class = if_else(class_2019 == "Dark", "Dark", "Light"))
kinase_entrez_lookup <- combined_pairs %>%
  distinct(Kinase, Kinase_entrez) %>%
  mutate(Kinase_entrez = as.character(Kinase_entrez))
stopifnot(!anyDuplicated(kinase_entrez_lookup$Kinase))   # one Entrez ID per symbol
dark_frac <- function(pairs) {
  d <- pairs %>%
    left_join(kinase_entrez_lookup, by = "Kinase") %>%
    inner_join(kin_class_final, by = "Kinase_entrez")
  100 * mean(d$Kinase_class == "Dark")
}

# The dark-kinase share is computed in two places (Figure 3B and the manuscript
# numbers). They must agree; a mismatch means a join key has drifted.
stopifnot(abs(100 * dark_fracs$frac_dark[dark_fracs$Set == "Curated + Specificity"] -
                dark_frac(distinct(bind_rows(pairs_curated, pairs_atlas)))) < 0.05)

manuscript_numbers <- tibble::tribble(
  ~quantity, ~value,
  "motif_percentile_cutoff",        as.character(motif_percentile_cutoff),
  "functional_score_cutoff",        as.character(functional_score_cutoff),
  "atlas_kinase_names_resolved",    as.character(nrow(atlas_kinase_map)),
  "atlas_kinase_genes",             as.character(n_distinct(atlas_kinase_map$Kinase)),
  "kinases_total",                  as.character(n_distinct(combined_pairs$Kinase)),
  "substrates_total",               as.character(n_distinct(combined_pairs$Substrate)),
  "phosphosites_total",             as.character(nrow(distinct(combined_pairs, Substrate, Site))),
  "pairs_curated",                  as.character(nrow(pairs_curated)),
  "pairs_atlas",                    as.character(nrow(pairs_atlas)),
  "pairs_novel",                    as.character(nrow(pairs_novel)),
  "pairs_total",                    as.character(nrow(distinct(bind_rows(pairs_curated, pairs_atlas)))),
  "pct_increase_over_curated",      sprintf("%.1f", 100 * nrow(pairs_novel) / nrow(pairs_curated)),
  "median_novel_substrates_per_kinase", as.character(median(novel_per_kinase$n_novel)),
  "max_novel_substrates_per_kinase",    as.character(max(novel_per_kinase$n_novel)),
  # Median over kinases of the fraction of a kinase's predicted substrates that share
  # a compartment with it. This is the quantity plotted in Figure 2A and it is the
  # fraction the co-localization filter retains: substrates with no HPA annotation
  # count as non-co-localized, exactly as in the filter itself.
  "median_pct_coloc_per_kinase", sprintf("%.1f", median(per_kinase_coloc$pct_coloc_all, na.rm = TRUE)),
  "pct_pass_S",                     sprintf("%.1f", residue_pass$pct_pass[residue_pass$residue == "S"]),
  "pct_pass_T",                     sprintf("%.1f", residue_pass$pct_pass[residue_pass$residue == "T"]),
  "pct_pass_Y",                     sprintf("%.1f", residue_pass$pct_pass[residue_pass$residue == "Y"]),
  "pct_func_only_S",                sprintf("%.1f", residue_pass$pct_func_only[residue_pass$residue == "S"]),
  "pct_func_only_T",                sprintf("%.1f", residue_pass$pct_func_only[residue_pass$residue == "T"]),
  "pct_func_only_Y",                sprintf("%.1f", residue_pass$pct_func_only[residue_pass$residue == "Y"]),
  "n_sites_S",                      as.character(residue_pass$n_sites[residue_pass$residue == "S"]),
  "n_sites_T",                      as.character(residue_pass$n_sites[residue_pass$residue == "T"]),
  "n_sites_Y",                      as.character(residue_pass$n_sites[residue_pass$residue == "Y"]),
  "dark_pct_curated",               sprintf("%.1f", dark_frac(pairs_curated)),
  "dark_pct_integrated",            sprintf("%.1f", dark_frac(distinct(bind_rows(pairs_curated, pairs_atlas))))
)

# chi-square test for the residue-class difference the manuscript reports
residue_tab <- cantley_sites90 %>%
  mutate(residue = substr(Site, 1, 1)) %>%
  filter(residue %in% c("S", "T", "Y")) %>%
  distinct(Substrate_Uniprot_ID, Site, residue, curated_site, ochoa_high) %>%
  mutate(pass = curated_site | ochoa_high) %>%
  count(residue, pass) %>%
  rename(ochoa_high = pass) %>%
  tidyr::pivot_wider(names_from = ochoa_high, values_from = n, values_fill = 0) %>%
  tibble::column_to_rownames("residue") %>%
  as.matrix()
residue_chisq <- chisq.test(residue_tab)
manuscript_numbers <- bind_rows(manuscript_numbers, tibble::tibble(
  quantity = c("residue_chisq_statistic", "residue_chisq_df", "residue_chisq_p"),
  value = c(sprintf("%.1f", unname(residue_chisq$statistic)),
            as.character(unname(residue_chisq$parameter)),
            format.pval(residue_chisq$p.value, digits = 3))
))

write_csv(manuscript_numbers, here("results/manuscript_numbers.csv"))
print(as.data.frame(manuscript_numbers), row.names = FALSE)
```

## Supplemental datasets

```{r}
# Supplementary Table S1: the complete unfiltered specificity-atlas prediction
# matrix, retaining continuous percentile, rank and promiscuity values.
write_csv(cantley_feats, here("results/unfiltered_cantley_dataset.csv"))

# Supplementary Table S2: gene-level HPA subcellular compartment assignments.
write_csv(subcellular_localization, here("results/subcellular_localization_mapping.csv"))

# Supplementary Table S3: Ochoa et al. phosphosite functional scores mapped onto
# our site identifiers.
write_csv(ochoa_fs_processed, here("results/ochoa_functional_scores.csv"))

# Relaxed motif tier: same contextual criteria, motif percentile lowered to
# `relaxed_percentile_cutoff`. Released separately and explicitly labelled so it
# cannot be mistaken for the high-confidence set.
relaxed_candidates <- cantley_feats %>%
  filter(!is.na(Percentile),
         Percentile >= relaxed_percentile_cutoff,
         Percentile <  motif_percentile_cutoff,
         promiscuity_index < promiscuity_index_cutoff,
         !is.na(Kinase), !is.na(Substrate)) %>%
  mutate(position = as.integer(position)) %>%
  left_join(ochoa_fs_norm, by = c("Substrate_Uniprot_ID", "position")) %>%
  filter(!is.na(functional_score), functional_score >= functional_score_cutoff)

# Co-localization is recomputed for this tier rather than reused: the top-tier
# counts were restricted to genes appearing in the stringent set, so reusing them
# would silently mark most relaxed-tier pairs as non-co-localized.
relaxed_genes <- unique(c(relaxed_candidates$Kinase_entrez,
                          relaxed_candidates$Substrate_entrez))
relaxed_hpa <- hpa_long %>% filter(Gene_entrez %in% relaxed_genes)
relaxed_coloc <- inner_join(
    relaxed_hpa %>% transmute(locs, Kinase_entrez = Gene_entrez) %>% distinct(),
    relaxed_hpa %>% transmute(locs, Substrate_entrez = Gene_entrez) %>% distinct(),
    by = "locs", relationship = "many-to-many"
  ) %>%
  count(Kinase_entrez, Substrate_entrez, name = "n_shared_compartments")

relaxed_tier <- relaxed_candidates %>%
  left_join(relaxed_coloc, by = c("Kinase_entrez", "Substrate_entrez")) %>%
  mutate(n_shared_compartments = coalesce(n_shared_compartments, 0L)) %>%
  filter(n_shared_compartments > 0L) %>%
  transmute(Kinase, Kinase_entrez, Substrate, Substrate_entrez,
            Substrate_Uniprot_ID, Site,
            Motif_percentile = Percentile, Motif_rank = Rank,
            Promiscuity_index = promiscuity_index,
            Functional_score = functional_score,
            Shared_compartments = n_shared_compartments,
            Tier = paste0("Relaxed (motif percentile ", relaxed_percentile_cutoff, "-",
                          motif_percentile_cutoff, ")")) %>%
  distinct()

write_csv(relaxed_tier, here("results/relaxed_tier_kinase_substrate_pairs.csv"))
message("Relaxed tier: ", nrow(relaxed_tier), " kinase-substrate-site rows across ",
        dplyr::n_distinct(paste(relaxed_tier$Kinase, relaxed_tier$Substrate)), " pairs.")
```

## Add in substrate-substrate interactions from HIPPIE

```{r, eval=FALSE}
## Adding in HIPPIE data for inter-substrate PPIs
#
# NOTE: this section is not part of the KinaseCanvas resource described in the
# manuscript; it builds a substrate-substrate PPI extension used by a separate
# downstream project. It is disabled because the HIPPIE file currently in
# `data/` is the native six-column HIPPIE export, not the MITAB release whose
# column names this code parses. Re-enable it only alongside a matching file.
# Read in Hippie data for PPIs
hippie <- read_tsv(here("data/hippie_current.txt"))

hippie  %>% 
  names()

hippie  %>% 
  head()  %>% 
  view()  

#Values in INteraction type and INteraction Identifer columns 
types = hippie$`Interaction Detection Methods`  %>% unique()


# Filter to only physical interactions between substrates

substrate_IDs <- unique(combined_pairs$Substrate_entrez)
kinase_IDs    <- unique(combined_pairs$Kinase_entrez)

substrate_IDs <- setdiff(substrate_IDs, kinase_IDs)
combined_IDs  <- union(substrate_IDs, kinase_IDs)

# --- HIPPIE → substrate-substrate + kinase autophosphorylation only ---
hippie_processed <- hippie %>% 
  select(
    Protein1_entrez = `ID Interactor A`,
    Protein2_entrez = `ID Interactor B`,
    confidence_score = `Confidence Value`,
    Interaction_detection_method = `Interaction Detection Methods`
  ) %>%
  mutate(
    Protein1_entrez = str_replace(Protein1_entrez, "^entrez gene:", ""),
    Protein2_entrez = str_replace(Protein2_entrez, "^entrez gene:", "")
  ) %>%
  # Node filter 
  filter(Protein1_entrez %in% combined_IDs,
         Protein2_entrez %in% combined_IDs) %>%
  distinct() %>%
  # classify edge type
  mutate(
    is_sub1 = Protein1_entrez %in% substrate_IDs,
    is_sub2 = Protein2_entrez %in% substrate_IDs,
    EdgeType = if_else(
      is_sub1 & is_sub2, "substrate-substrate", "exclude"
    )
  ) %>%
  # drop everything else:
  #  - kinase–substrate (is_kinX & is_subY)
  #  - kinase–kinase (kinases)
  #  - self-loops (if present in HIPPIE; rare but possible)
  #  - substrate autoloops (if present in HIPPIE; usually rare/artefactual)
  filter(EdgeType != "exclude") %>%
  select(Protein1_entrez, Protein2_entrez, confidence_score,
         Interaction_detection_method, EdgeType) %>%
  distinct()

hippie_processed  %>% head()  %>% view()

# #quick look at confidence score distribution
# ggplot(hippie_processed, aes(confidence_score)) +
#   geom_histogram(bins=100, fill="grey70", color="black") +
#   #lines for quantiles
#   geom_vline(xintercept = quantile(hippie_processed$confidence_score, probs=c(0.75, 0.9)), linetype="dashed", color="red") +
#   labs(x="Confidence score", y="Count") +
#   scale_y_continuous(labels=comma, expand=expansion(mult=c(0,0.1))) +  # 10% headroom
#   theme_pubr()

confidence_cutoff = quantile(hippie_processed$confidence_score, probs=0.75)

# #quick look at MI codes in Interaction Detection Methods column (need to split by "|")
# #also show Ns in the dataset
# hippie %>%
#   filter(`Confidence Value` > confidence_cutoff) %>%
#   select(`Interaction Detection Methods`) %>%
#   separate_rows(`Interaction Detection Methods`, sep="\\|") %>%
#   group_by(`Interaction Detection Methods`)  %>%
#   summarise(N = n())  %>%
#   arrange(desc(N))  %>%
#   ungroup()  %>%
#   mutate(`Interaction Detection Methods` = str_replace_all(`Interaction Detection Methods`, "MI:", ""))  %>%
#   arrange(desc(N))  %>%
#   view()
# ## Assays to filter out: 
# 0428(imaging technique)
# 0493(in vivo)
# 0492(in vitro)

hippie_symbol_to_entrez_1 = bitr(hippie_processed$Protein1_entrez, fromType="ENTREZID", toType=c("SYMBOL"), OrgDb="org.Hs.eg.db")  %>%
  rename(Protein1_entrez = ENTREZID, Protein1 = SYMBOL)  %>%
  distinct()

hippie_symbol_to_entrez_2 = bitr(hippie_processed$Protein2_entrez, fromType="ENTREZID", toType=c("SYMBOL"), OrgDb="org.Hs.eg.db")  %>%
  rename(Protein2_entrez = ENTREZID, Protein2 = SYMBOL)  %>%
  distinct()

hippie_entrez = hippie_processed  %>%
  left_join(hippie_symbol_to_entrez_1, by="Protein1_entrez")  %>%
  left_join(hippie_symbol_to_entrez_2, by="Protein2_entrez")  %>%
  drop_na(Protein1, Protein2)  %>%
  distinct() %>%
  # Consulted with proteomics colleagues, decided to filter out certain methods
  filter(Interaction_detection_method != "MI:0428")  %>% #remove imaging technique (too vague)
  filter(Interaction_detection_method != "MI:0493")  %>% #remove in vivo (too vague)
  filter(Interaction_detection_method != "MI:0492")  %>% #remove in vitro (too vague)
  filter(Protein1_entrez != Protein2_entrez)  %>% #remove self-loops
  filter(confidence_score > confidence_cutoff)  %>%
  select(Protein1, Protein2, Protein1_entrez, Protein2_entrez, Evidence = Interaction_detection_method)  %>%
  mutate(Source = "HIPPIE")  %>%
  distinct()


# Combine with integrated data
full_interaction_data = combined_pairs  %>%
  select(
    Protein1 = Kinase,
    Protein2 = Substrate,
    Protein1_entrez = Kinase_entrez,
    Protein2_entrez = Substrate_entrez,
    Evidence,
    Source
  )  %>%
  mutate(Protein1_entrez = as.character(Protein1_entrez),
         Protein2_entrez = as.character(Protein2_entrez))  %>% 
  bind_rows(hippie_entrez)  %>%
  distinct() %>%
  write_csv(here("results/extended_kinome_network.csv"))

#summarise across sources
full_interaction_summary = full_interaction_data  %>%
  group_by(Protein1, Protein2, Protein1_entrez, Protein2_entrez)  %>%
  summarise(
    n_sources = n_distinct(Source),
    Sources   = paste(sort(unique(Source)), collapse="; "),
    Evidence  = paste(sort(unique(Evidence)), collapse="; "),
    .groups = "drop"
  ) %>%
  write_csv(here("results/extended_kinome_network_summary.csv"))

```